---
title: "Gridded output Storm events (Wales 2018-2021)"
subtitle: "Sim & SUFI data"
author: "Caleb Scutt"
format:
  html:
    page-layout: full
    toc: true
    code-fold: true
    fig-width: 10
    fig-height: 5
    embed-resources: true
    theme: flatly
    css: styles.css
  typst:
    toc: true
    fig-width: 7
    fig-height: 5
    margin:
      x: 1in
      y: 1in
    fontsize: 11pt
    section-numbering: 1.1.1
    keep-typ: false
execute:
  warning: false
  message: false
---


## Load functions and packages

```{r}
library(data.table)
library(ggplot2)
library(lubridate)
library(gtools)
library(sf)
library(grid)
library(gridExtra)
library(patchwork)
library(gt)
library(terra)
library(elevatr)
source("../functions/load.g2g.runs.R")
source("../functions/extract.g2g.grid.R")
# source("../functions/qt.exceed.R")
source("../functions/plot.wales.R")
# source("../functions/obs.mod.stats.R")
source("../functions/qt.exceed.R")
source("../functions/plot.wales.R")
source("../functions/obs.mod.stats.R")
#
```


## Read in gridded output

```{r}

# g2g_output_2018_2022 <- load.g2g.data(dates = c(2018, 2021), sufi = TRUE)

events <- fread("../data/notable_events.csv")
events_pre_2022 <- events[Year <= 2021,]
j <- 0
storm_names <- c()
for (i in 1:nrow(events_pre_2022)){
 event <- events_pre_2022[i, ]
        if (!is.na(event[, Storm_name])){
            storm_name <- event[, Storm_name]
        } else {
            j <- j + 1
            storm_name <- paste0("Unnamed_storm_", j) ## assign unnamed storms a name for reference to csv
        }
        storm_names[i] <- storm_name
}

## read in grids
sim_path <- "../data/g2g_data/flow_grids/maxFlows_Sim"
sufi_path <- "../data/g2g_data/flow_grids/maxFlows_SUFI"
paths <- list(sim = sim_path, sufi =  sufi_path)
grid_files <- lapply(paths, list.files, include.dirs = TRUE)
grid_file_paths <- Map(function(path, file){
    file.path(path, file)
}, paths, grid_files)


g2g_event_grids <- lapply(grid_file_paths, lapply, load.maxflow.event)
g2g_event_grids <- lapply(g2g_event_grids, function(x) {names(x) <- storm_names
return(x)})
# g2g_output_2018_2022 <- load.g2g.data(dates = c(2018, 2021), sufi = TRUE)
# obs_ts <- g2g_output_2018_2022$obs
# events_list <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  NULL, exclude.non.notable.sites = FALSE)

# g2g_event_grids <- Map(load.maxflow.event(file.path(paths[[1]],grid_files)[1:9]), grid_files[[1]])



```

## Read in gridded QT

```{r}
## read qt grids
qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grids <- lapply(qt_grid_paths, function(f) normalise.to.corner(read.ascii(f)))
names(qt_grids) <- qt_val
check.grid.alignment(qt_grids) ## sanitise check
qt_rasts <- lapply(qt_grids, ascii.to.rast) ## convert to terra format
names(qt_rasts) <- qt_val ## name by qt val
qt_order <- c("qmed", "q5", "q10", "q25", "q50", "q100", "q250", "q1000")
qt_rasts <- qt_rasts[intersect(qt_order, names(qt_rasts))]

wales <- st_read("../data/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")
wales <- st_transform(wales, 27700)

catchments <- fread("../data/cdata/catchment_cdata_EA-NRW.csv", fill = Inf)


catchments_sf <- st_as_sf(
  catchments,
  coords = c("WISKI.EASTING", "WISKI.NORTHING"),
  crs = 27700,
  remove = FALSE
)
wales_poly <- st_union(wales)
sf_g2g_ids <- unique(c(catchments_sf[st_intersects(catchments_sf, wales_poly, sparse = FALSE), ]$G2G.ID, catchments[Region. == "Wales"]$G2G.ID)) ## use both ids coded as Wales, plus anything inside the polygon
wales_cdata <- catchments[G2G.ID %in% sf_g2g_ids]## filter by wales


```


## cutout shape file

```{r}
# # Catchment polygons (all regions)
# catchments_sf <- st_read("../data/20250610_g2g_ffc_v2_0/fullcats_g2g_ffc_v2_0_loc.shp")

# # Wales boundary polygon
# wales <- st_read("../data/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")

# # Match CRS
# wales <- st_transform(wales, st_crs(catchments_sf))

# # Dissolve to one Wales geometry (prevents duplicate overlaps)
# wales_poly <- st_union(wales)

# # Keep only catchments that intersect Wales
# catchments_wales <- catchments_sf[st_within(catchments_sf, wales_poly, sparse = FALSE), ]
# catchments_wales <- catchments_sf[st_within(catchments_sf, wales_poly, sparse = FALSE), ]
# catchment_wales <- catchments_sf[catchments_sf$g2g_id %in% wales_cdata$G2G.ID, ]
# # Optional: write out
# dir.create("../data/wales_catchments_sf", showWarnings = FALSE, recursive = TRUE)

# st_write(
#   catchment_wales,
#   dsn = "../data/wales_catchments_sf",
#   layer = "wales_catchments_sf",
#   driver = "ESRI Shapefile",
#   delete_layer = TRUE
# )

wales_catchment_sf <- st_read("../data/wales_catchments_sf")


```

```{r}
g2g_output_2018_2022 <- load.g2g.data(dates = c(2018, 2021), sufi = TRUE)
obs_ts <- g2g_output_2018_2022$obs
mod_ts <- g2g_output_2018_2022$mod

events_list <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  NULL, exclude.non.notable.sites = FALSE)

## change dates so it fits with gridded dates
ciara_pstart <-  as.POSIXct(paste(2020, 2, 1, sep="-"), tz="GMT")
ciara_pend <-  as.POSIXct(paste(2020, 2, 16, sep="-"), tz="GMT")
dennis_pstart <-  as.POSIXct(paste(2020, 2, 15, sep="-"), tz="GMT")
dennis_pend <-  as.POSIXct(paste(2020, 2, 22, sep="-"), tz="GMT")
unnamed_2_pstart <-  as.POSIXct(paste(2020, 2, 21, sep="-"), tz="GMT")
unnamed_2_pend <-  as.POSIXct(paste(2020, 2, 28, sep="-"), tz="GMT")
jorge_pstart <-  as.POSIXct(paste(2020, 2, 27, sep="-"), tz="GMT")
jorge_pend <-  as.POSIXct(paste(2020, 3, 9, sep="-"), tz="GMT")

events_list <- lapply(events_list, function(config){

  config$Ciara$obs <- config$Ciara$obs[DATE_TIME >= ciara_pstart &
                     DATE_TIME <= ciara_pend, ]

  config$Dennis$obs <- config$Dennis$obs[DATE_TIME >= dennis_pstart &
                      DATE_TIME <= dennis_pend, ]

  config$Unnamed_storm_2$obs <- config$Unnamed_storm_2$obs[DATE_TIME >= unnamed_2_pstart &
                         DATE_TIME <= unnamed_2_pend, ]

  config$Jorge$obs <- config$Jorge$obs[DATE_TIME >= jorge_pstart &
                     DATE_TIME <= jorge_pend, ]

  config
})


qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grid_list <- lapply(qt_grid_paths, read.ascii)
names(qt_grid_list) <- qt_val ## name by qt
qt_dt <- make.qt.csv(qt_grid_list, cdata = wales_cdata)

sim_max_discharge_qt <- list()
for (t_val in qt_val) {
  sim_max_discharge_qt[[t_val]] <- lapply(events_list$sim, function(event_mod_obs) {
    extract.peak.discharge(event_mod_obs, qt_dt, t_val)
  })
}

sufi_max_discharge_qt <- list()
for (t_val in qt_val) {
  sufi_max_discharge_qt[[t_val]] <- lapply(events_list$sufi, function(event_mod_obs) {
    extract.peak.discharge(event_mod_obs, qt_dt, t_val)
  })
}

max_discharge_qt <- list(sim = sim_max_discharge_qt, sufi = sufi_max_discharge_qt)

qt_exceeded <- lapply(max_discharge_qt,lapply, lapply, thresh.exceed)


qt_exceeded_dt_list <- lapply(qt_exceeded, lapply, function(qt){
    dt_exceed <- data.table(
    unsplit_name     = names(unlist(qt)),
    Exceeded = unlist(qt)
    )
    dt_exceed[, c("Storm", "Data_Type", "G2G.ID") := tstrsplit(unsplit_name, "\\.", keep = 1:3)]
    dt_exceed[, G2G.ID := gsub("_Obs|_Mod", "", G2G.ID)]
    dt_exceed[, unsplit_name := NULL]
    setcolorder(dt_exceed, c("Storm", "Data_Type", "G2G.ID", "Exceeded"))
    return(dt_exceed)
})

merged_dat <- lapply(qt_exceeded_dt_list, lapply, merge, qt_dt, by = "G2G.ID")
q_levels <- c("q1", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")

highest_qt_exceeded_list <- lapply(merged_dat, function(model_config){
    exceed_dt <- rbindlist(model_config, idcol = "Threshold")

    exceed_dt[, Threshold := gsub("qmed", "q1", Threshold)] ## sub qmed for q1 so it can be ordered correctly in next step

    exceed_dt[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

    exceed_dt[, keep_site_storm := any(Exceeded == TRUE, na.rm = TRUE), by = .(Storm, G2G.ID)] ## identify sites exceeded in any of 3 data (obs, sim sufi)
    filtered_dt <- exceed_dt[keep_site_storm == TRUE] ## filter these sites

    highest_qt_exceeded <- filtered_dt[, .(
      Threshold = if (any(Exceeded == TRUE, na.rm = TRUE)) {
        as.character(max(Threshold[Exceeded == TRUE], na.rm = TRUE))
      } else {
        "None"
      },
      Site         = first(Site),
      G2G_Easting  = first(G2G_Easting),
      G2G_Northing = first(G2G_Northing)
    ), by = .(Storm, G2G.ID, Data_Type)]

    highest_qt_exceeded[, Threshold := gsub("^q1$", "qmed", Threshold)]

    final_levels <- c("None", "qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")
    highest_qt_exceeded[, Threshold := factor(Threshold, levels = final_levels, ordered = TRUE)]

    # Optional: Sort it cleanly for viewing
    setorder(highest_qt_exceeded, Storm, G2G.ID, Data_Type)
    return(highest_qt_exceeded)
})

q_levels <- c("qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000") ## reassign q_levels back to normal


```




## Plot

```{r}
#| fig-width: 20
#| fig-height: 7

catchments <- fread("../data/cdata/catchment_cdata_EA-NRW.csv", fill = Inf)
## filter by wales
wales <- st_read("../data/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")
wales <- st_transform(wales, 27700)

catchments_sf <- st_as_sf(
  catchments,
  coords = c("WISKI.EASTING", "WISKI.NORTHING"),
  crs = 27700,
  remove = FALSE
)
wales_poly <- st_union(wales)
sf_g2g_ids <- unique(c(catchments_sf[st_intersects(catchments_sf, wales_poly, sparse = FALSE), ]$G2G.ID, catchments[Region. == "Wales"]$G2G.ID)) ## use both ids coded as Wales, plus anything inside the polygon
wales_cdata <- catchments[G2G.ID %in% sf_g2g_ids]## filter by wales

g2g_data <- wales_cdata[, .(G2G.ID, G2G.Easting, G2G.Northing, CATCHMENTSIZE)]

event_exceedance_grids <- lapply(g2g_event_grids, lapply, build.exceedance.grid, qt_rasts)

for (storm in storm_names){
    p <- plot.gridded.exceedance.map(event_exceedance_grids[["sim"]][[storm]], wales, event_label = storm, g2g.sites = NULL, catchments_sf = wales_catchment_sf)  + labs(title = paste(" Sim QT gridded exceedance -", storm), x = NULL, y = NULL)

    q <- plot.gridded.exceedance.map(event_exceedance_grids[["sufi"]][[storm]], wales, event_label = storm, g2g.sites = NULL, catchments_sf = wales_catchment_sf)+ labs(title = paste(" Sufi QT gridded exceedance -", storm), x = NULL, y = NULL)
    r <- plot.wales.exceed(highest_qt_exceeded_list$sufi, wales, storm.name = storm, "obs")

    print(p + q + r)
}





```


## FOCUS ON SPECIFIC CATCHMENTS?

### Elwy at Ciara

```{r}
ciara_exceedance_grid_sufi <- event_exceedance_grids$sufi$Ciara
ciara_exceedance_grid_sim <- event_exceedance_grids$sim$Ciara

catchments <- st_read("../data/20250610_g2g_ffc_v2_0/fullcats_g2g_ffc_v2_0_loc.shp")
elwy_catchment_id <- wales_cdata[wales_cdata$River.Name == "ELWY[N][WA]"]$G2G.ID
elwy_catchment_sf <- catchments[catchments$g2g_id == elwy_catchment_id,]


# target_catchment <- st_transform(target_catchment, st_crs(g2g_rast))

rast_cropped_sufi <- terra::crop(ciara_exceedance_grid_sufi, elwy_catchment_sf)
rast_cropped_sim <- terra::crop(ciara_exceedance_grid_sim, elwy_catchment_sf)


rast_cutout_sim <- terra::mask(rast_cropped_sim, elwy_catchment_sf)
rast_cutout_sufi <- terra::mask(rast_cropped_sufi, elwy_catchment_sf)

df_cutout_sim <- as.data.frame(rast_cutout_sim, xy = TRUE)
df_cutout_sufi <- as.data.frame(rast_cutout_sufi, xy = TRUE)


names(df_cutout_sim)[3] <- "value" 
names(df_cutout_sufi)[3] <- "value" 


threshold_colours <- c(
    "None"  = "white",
    "qmed"  = "#08306b",
    "q5"    = "#2171b5",
    "q10"   = "#41b6c4",
    "q25"   = "#ffff33",
    "q50"   = "#fd8d3c",
    "q100"  = "#e31a1c",
    "q250"  = "#99000d",
    "q1000" = "#4d0000"
  )

  
ggplot() +
  geom_raster(data = df_cutout_sim, aes(x = x, y = y, fill = value)) +
  
  geom_sf(data = elwy_catchment_sf, fill = NA, colour = "black", linewidth = 0.6) +
  

  coord_sf(datum = st_crs(27700)) +
  labs(title = "sim G2G Cutout: Elwy Catchment", x = NULL, y = NULL) +
  theme_minimal() +
  scale_fill_manual(
      values = threshold_colours, 
      breaks = names(threshold_colours), # Removed the trailing comma here
      name = "QT Threshold exceeded", 
      drop = FALSE, 
      na.value = NA
    ) 



ggplot() +
  geom_raster(data = df_cutout_sufi, aes(x = x, y = y, fill = value)) +
  
  geom_sf(data = elwy_catchment_sf, fill = NA, colour = "black", linewidth = 0.6) +
  

  coord_sf(datum = st_crs(27700)) +
  labs(title = "SUFI G2G Cutout: Elwy Catchment", x = NULL, y = NULL) +
  theme_minimal() +
  scale_fill_manual(
      values = threshold_colours, 
      breaks = names(threshold_colours), # Removed the trailing comma here
      name = "QT Threshold exceeded", 
      drop = FALSE, 
      na.value = NA
    ) 


```


## Extract topography and calculate TRI

```{r}
# all_zips <- list.files(
#   path = "../data/topography/os_raw",
#   pattern = "\\.zip$",
#   recursive = TRUE,
#   full.names = TRUE
# )

# # (tolower makes sure we capture it regardless of upper/lower case names)
# wales_prefixes <- "/(sh|sj|sm|sn|so|ss|st)/"
# welsh_zips <- all_zips[grepl(wales_prefixes, tolower(all_zips))]

# for (zip_file in welsh_zips) {
#   unzip(zip_file, exdir = dirname(zip_file))
# }

# all_tiles <- list.files(
#   path = "../data/topography/os_raw", 
#   pattern = "\\.asc$", 
#   recursive = TRUE, 
#   ignore.case = TRUE,
#   full.names = TRUE
# )

# # keep welsh tiles
# wales_tile_paths <- all_tiles[grepl(wales_prefixes, tolower(all_tiles))]

# print(length(wales_tile_paths))

# dem_virtual <- terra::vrt(wales_tile_paths)

# crs(dem_virtual) <- "EPSG:27700"

# dem_virtual <- terra::vrt(wales_tile_paths)
# crs(dem_virtual) <- "EPSG:27700"

# wales_dem <- terra::crop(dem_virtual, wales)
# wales_dem <- terra::mask(wales_dem, wales)

# writeRaster(wales_dem, "../data/topography/wales_dem_50m.tif", overwrite = TRUE)
# wales_dem <- terra::rast("../data/topography/wales_dem_50m.tif")

# slope_rast <- terra::terrain(wales_dem, v = "slope", unit = "degrees")
# tri_rast   <- terra::terrain(wales_dem, v = "TRI")


# catchments_v <- terra::vect(catchments)

# print("Extracting metrics for each catchment...")
# slope_extracted <- terra::extract(slope_rast, catchments_v, fun = mean, na.rm = TRUE)
# tri_extracted   <- terra::extract(tri_rast, catchments_v, fun = mean, na.rm = TRUE)
# tri_range_extracted   <- terra::extract(tri_rast, catchments_v, fun = sd, na.rm = TRUE) ## range might be more interesting for heterogeneity ?

# terrain_dt <- data.table(
#   G2G.ID     = catchments$g2g_id, 
#   Mean_Slope = slope_extracted[[2]],  
#   Mean_TRI   = tri_extracted[[2]]
# )

# wales_cdata <- merge(wales_cdata, terrain_dt, by = "G2G.ID", all.x = TRUE)

# head(wales_cdata[, .(G2G.ID, River.Name, Mean_Slope, Mean_TRI)])

```

## TODO
- add the catchment polygons
- export shape file of just wales catchments




## RAINGAUGE DENSITY METRICS

```{r}



```

## next things to do

- overlay g2g sites?
- overlay raingauge ?
- topography 
  - downloaded OS map of topography.

