---
title: "QT Exceedannce plots  for Storm events (Wales 2018-2021)"
subtitle: "Sim & SUFI data"
author: "Caleb Scutt"
format:
  html:
    page-layout: full
    toc: true
    code-fold: true
    fig-width: 10
    fig-height: 5
    embed-resources: true
    theme: flatly
  typst:
    toc: true
    fig-width: 7
    fig-height: 5
    margin:
      x: 1in
      y: 1in
    fontsize: 11pt
    section-numbering: 1.1.1
    keep-typ: false
execute:
  warning: false
  message: false
---


```{r echo = FALSE}
## load packages and functions
library(data.table)
library(ggplot2)
library(lubridate)
library(gtools)
library(sf)
library(grid)
library(gridExtra)
library(patchwork)
library(gt)
library(terra)
library(lme4)
source("../functions/load.g2g.runs.R")
source("../functions/extract.g2g.grid.R")
# source("../functions/qt.exceed.R")
source("../functions/plot.wales.R")
source("../functions/obs.mod.stats.R")
```


<!-- ## Read in data

* `mod_ts` is the g2g modelled flow data at each g2g site
* `obs_ts` is the observed flow data at each g2g site
* `events` are the notable storm events in Wales (`events_pre_2022` are events we have data for)
* `catchments` are the metadata for each catchment (region, area, river name, easting/northing, g2g.id). `wales_cdata` is this filtered by just Welsh catchments
* The following sites are missing from g2g model output, but present in catchment data: 055025_TG_325_Mod, 57809_Mod 60005_Mod, 066004_TG_9512_Mod, 067005_TG_116_Mod, 067023_TG_117_Mod, 055025_TG_325_Mod, 60005_Mod. -->

```{r, include = FALSE}
g2g_output_2018_2022 <- load.g2g.data(dates = c(2018, 2021), sufi = TRUE)
obs_ts <- g2g_output_2018_2022$obs
mod_ts <- g2g_output_2018_2022$mod
## read in the events
events <- fread("../data/notable_events.csv")
events_pre_2022 <- events[Year <= 2021,]
## read in the catchment metadata
catchments <- fread("../data/cdata/catchment_cdata_EA-NRW.csv", fill = Inf)
## filter by wales
wales <- st_read("../data/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")
wales <- st_transform(wales, 27700)

catchments_sf <- st_as_sf(
  catchments,
  coords = c("WISKI.EASTING", "WISKI.NORTHING"),
  crs = 27700,
  remove = FALSE
)
wales_poly <- st_union(wales)
sf_g2g_ids <- unique(c(catchments_sf[st_intersects(catchments_sf, wales_poly, sparse = FALSE), ]$G2G.ID, catchments[Region. == "Wales"]$G2G.ID)) ## use both ids coded as Wales, plus anything inside the polygon
wales_cdata <- catchments[G2G.ID %in% sf_g2g_ids]## filter by wales


```


<!-- Here we make a nested list, with top level of list representing each storm event from `events`. <br>
We subset Wales into North, Mid and South by these Northing values (can be adjusted). These were used because some sites were not categorised by Area (e.g South East [CY]) <br>
We subset the time series 7 days either side of the event beginning/end. <br> -->


```{r message = FALSE, warning = FALSE, include = FALSE, echo = TRUE}

events_list <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  NULL, exclude.non.notable.sites = FALSE)

```


<!-- ## Read in QT data

QT data is extracted using functions from `functions/griddedData_class_jrw.py` converted to R.
These functions can be re-used when more catchment data is added.
Note that these QT values are calculated pre-2022. -->



```{r, message = FALSE, include = FALSE}
## this chunk extracts the g2g qt vals
qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grid_list <- lapply(qt_grid_paths, read.ascii)
names(qt_grid_list) <- qt_val ## name by qt
qt_dt <- make.qt.csv(qt_grid_list, cdata = wales_cdata, write = TRUE)
```


## Extract the peak discharge at each flood event

* Peak discharge extracted at each event at each g2g site for observed (what actually happened) and modelled (sim and sufi).

```{r, include = FALSE}

sim_max_discharge_qt <- list()
for (t_val in qt_val) {
  sim_max_discharge_qt[[t_val]] <- lapply(events_list$sim, function(event_mod_obs) {
    extract.peak.discharge(event_mod_obs, qt_dt, t_val)
  })
}

sufi_max_discharge_qt <- list()
for (t_val in qt_val) {
  sufi_max_discharge_qt[[t_val]] <- lapply(events_list$sufi, function(event_mod_obs) {
    extract.peak.discharge(event_mod_obs, qt_dt, t_val)
  })
}

max_discharge_qt <- list(sim = sim_max_discharge_qt, sufi = sufi_max_discharge_qt)

qt_exceeded <- lapply(max_discharge_qt,lapply, lapply, thresh.exceed)


qt_exceeded_dt_list <- lapply(qt_exceeded, lapply, function(qt){
    dt_exceed <- data.table(
    unsplit_name     = names(unlist(qt)),
    Exceeded = unlist(qt)
    )
    dt_exceed[, c("Storm", "Data_Type", "G2G.ID") := tstrsplit(unsplit_name, "\\.", keep = 1:3)]
    dt_exceed[, G2G.ID := gsub("_Obs|_Mod", "", G2G.ID)]
    dt_exceed[, unsplit_name := NULL]
    setcolorder(dt_exceed, c("Storm", "Data_Type", "G2G.ID", "Exceeded"))
    return(dt_exceed)
})

merged_dat <- lapply(qt_exceeded_dt_list, lapply, merge, qt_dt, by = "G2G.ID")

```


```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
# wales <- st_read("W:/hymod/Hydro-JULES/HJ Internships/2026 - NRW G2G/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")# Read Wales constituency polygons
# 

q_levels <- c("q1", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")

highest_qt_exceeded_list <- lapply(merged_dat, function(model_config){
    exceed_dt <- rbindlist(model_config, idcol = "Threshold")

    exceed_dt[, Threshold := gsub("qmed", "q1", Threshold)] ## sub qmed for q1 so it can be ordered correctly in next step

    exceed_dt[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

    exceed_dt[, keep_site_storm := any(Exceeded == TRUE, na.rm = TRUE), by = .(Storm, G2G.ID)] ## identify sites exceeded in any of 3 data (obs, sim sufi)
    filtered_dt <- exceed_dt[keep_site_storm == TRUE] ## filter these sites

    highest_qt_exceeded <- filtered_dt[, .(
      Threshold = if (any(Exceeded == TRUE, na.rm = TRUE)) {
        as.character(max(Threshold[Exceeded == TRUE], na.rm = TRUE))
      } else {
        "None"
      },
      Site         = first(Site),
      G2G_Easting  = first(G2G_Easting),
      G2G_Northing = first(G2G_Northing)
    ), by = .(Storm, G2G.ID, Data_Type)]

    highest_qt_exceeded[, Threshold := gsub("^q1$", "qmed", Threshold)]

    final_levels <- c("None", "qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")
    highest_qt_exceeded[, Threshold := factor(Threshold, levels = final_levels, ordered = TRUE)]

    # Optional: Sort it cleanly for viewing
    setorder(highest_qt_exceeded, Storm, G2G.ID, Data_Type)
    return(highest_qt_exceeded)
})

q_levels <- c("qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000") ## reassign q_levels back to normal


```

## QT Exceeded plots

* The following plots show the highest flow threshold exceeded at each site, for each storm.

```{r echo = FALSE}
#| fig-width: 14
#| fig-height: 5


for (name in unique(highest_qt_exceeded_list[[1]]$Storm)) {
  p <- plot.wales.exceed(highest_qt_exceeded_list$sufi, wales, storm.name = name, "obs")
  q <- plot.wales.exceed(highest_qt_exceeded_list$sim, wales, storm.name = name, "sim")
  r <- plot.wales.exceed(highest_qt_exceeded_list$sufi, wales, storm.name = name, "sufi")

  pqr <- list(obs = p, sim = q,  sufi = r)

  # lapply(names(pqr), function(config){

  #   plot <- pqr[[config]]

  #   ggsave(
  #         filename = file.path(paste0("../data/figures/qt_exceed_maps"), paste0(name, "_", config, "_qt_exceed_map", ".png")),
  #         plot = plot,
  #         width = 8,
  #         height = 6,
  #         units = "in",
  #         dpi = 300
  #       )
  # })

  print(
    (p + q + r) +
      plot_annotation(title = paste("Storm:", name))
  )
}

```



```{r echo = FALSE}
#| fig-width: 15
#| fig-height: 20


comp_dt_list <- lapply(highest_qt_exceeded_list, function(x) {
  comp_dt <- dcast(x, Storm + G2G.ID + Site + G2G_Easting + G2G_Northing ~ Data_Type, value.var = "Threshold" )
  comp_dt[, rank_mod := as.numeric(mod)] ## ranked where qmed is 1, q1000 is 10
  comp_dt[, rank_obs := as.numeric(obs)]
  comp_dt[is.na(rank_mod), rank_mod := 0]
  comp_dt[is.na(rank_obs), rank_obs := 0]
  comp_dt[, rank_diff := abs(rank_mod - rank_obs)] ## calculate rank differences
})


```

## Prediction accuracy metrics

* POD = probability of detection (closer to 1 = better)
* FAR = false alarm ratio (closer to 0 = better)
* CSI = critical success ratio  (closer to 1 = better)

```{r echo = TRUE}
#| fig-height: 5


# 1. Count the occurrences of each obs/mod combination
sim_heatmap_dt <- comp_dt_list$sim[obs != "None" & mod != "None", .N, by = .(obs, mod)]
sim_heatmap_dt[, `:=`(
  obs = droplevels(obs), ## this removed the unused "None " factor from the data.table
  mod = droplevels(mod)
)] ## 

# 2. Plot the heatmap
p <- ggplot(sim_heatmap_dt, aes(x = obs, y = mod, fill = N)) +
  geom_tile(color = "white", size = 0.5) + 
  geom_text(aes(label = N, color = N > max(N)/2), show.legend = FALSE, size = 4) +
  scale_color_manual(values = c("black", "white")) +
  
  # A clean gradient (white to steelblue works well for hydrology reports)
  scale_fill_distiller(palette = "Blues", direction = 1, name = "Count") +  # Ensure all factor levels display on the axes, even if the count is 0
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  
  labs(
    title = "Sim model vs Observed Discharge Thresholds",
    x = "Observed Threshold (obs)",
    y = "Modelled Threshold (SIM)"
  ) +
  
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Removes default grid lines for a cleaner look
    axis.text.x = element_text(angle = 45, hjust = 1) # Angles x-axis labels if they get crowded
)


sufi_heatmap_dt <- comp_dt_list$sufi[obs != "None" & mod != "None", .N, by = .(obs, mod)]
sufi_heatmap_dt[, `:=`(
  obs = droplevels(obs), ## this removed the unused "None " factor from the data.table
  mod = droplevels(mod)
)] ## 

# 2. Plot the heatmap
q <- ggplot(sufi_heatmap_dt, aes(x = obs, y = mod, fill = N)) +
  geom_tile(color = "white", size = 0.5) + 
   geom_text(aes(label = N, color = N > max(N)/2), show.legend = FALSE, size = 4) +
  scale_color_manual(values = c("black", "white")) +
  
  # A clean gradient (white to steelblue works well for hydrology reports)
  scale_fill_distiller(palette = "Blues", direction = 1, name = "Count") +  # Ensure all factor levels display on the axes, even if the count is 0
  # Ensure all factor levels display on the axes, even if the count is 0
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  
  labs(
    title = "SUFI model vs Observed Discharge Thresholds",
    x = "Observed Threshold (obs)",
    y = "Modelled Threshold (SUFI)"
  ) +
  
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Removes default grid lines for a cleaner look
    axis.text.x = element_text(angle = 45, hjust = 1) # Angles x-axis labels if they get crowded
)
p <- p +
  theme(
    plot.background  = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),

    axis.text.x  = element_text(colour = "white", size = 18, face = "bold"),
    axis.text.y  = element_text(colour = "white", size = 18, face = "bold"),
    axis.title.x = element_text(colour = "white"),
    axis.title.y = element_text(colour = "white"),
    axis.ticks   = element_line(colour = "white"),

    legend.text  = element_text(colour = "white"),
    legend.title = element_text(colour = "white"),

    plot.title    = element_text(colour = "white"),
    plot.subtitle = element_text(colour = "white"),
    plot.caption  = element_text(colour = "white")
  )

q <- q + theme(
    plot.background  = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),

    axis.text.x  = element_text(colour = "white", size = 18, face = "bold"),
    axis.text.y  = element_text(colour = "white", size = 18, face = "bold"),
    axis.title.x = element_text(colour = "white"),
    axis.title.y = element_text(colour = "white"),
    axis.ticks   = element_line(colour = "white"),

    legend.text  = element_text(colour = "white"),
    legend.title = element_text(colour = "white"),

    plot.title    = element_text(colour = "white"),
    plot.subtitle = element_text(colour = "white"),
    plot.caption  = element_text(colour = "white")
  )



# ggsave( ## save files here
#           filename = "../data/figures/qt_heatmap_sim.svg",
#           plot = p,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300,
#           bg = "transparent"
#         )

# ggsave(
#           filename = "../data/figures/qt_heatmap_sufi.svg",
#           plot = q,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300,
#           bg = "transparent"
#         )



p + q


```

## Peak magnitude error

All sites that are within regions affected for each storm event <br>

Note that the upper scale is cutoff at 98th quantile to prevent outliers controlling colour range. <br>

Grey sites = missing data. <br>

Positive = mod overestimated <br>

```{r echo = FALSE}
#| fig-width: 20
#| fig-height: 10
pme_list <- lapply(events_list, lapply, peak.magnitude.error)

all_pme_plots <- setNames(lapply(names(pme_list), function(mod_config) {
  sim_type <- mod_config

  pme_dt <- data.table(
    unsplit_name = names(unlist(pme_list[[mod_config]])),
    pme = unlist(pme_list[[mod_config]])
  )

  pme_dt[, c("Storm", "G2G.ID") := tstrsplit(unsplit_name, "\\.")]
  pme_dt[, unsplit_name := NULL]
  setcolorder(pme_dt, c("Storm", "G2G.ID", "pme"))

  merged_pme <- merge(pme_dt, qt_dt, by = "G2G.ID")[, .(G2G.ID, Storm, pme, G2G_Easting, G2G_Northing)]

  storm_names <- unique(merged_pme$Storm)

  # Wrap the inner lapply in setNames() so each plot is named by its Storm
  p_list <- setNames(lapply(storm_names, function(name) {
    p <- plot.wales.pme(merged_pme, wales, storm.name = name)
    
    # Optional: Add a subtitle to each plot so you know which model is which when combined
    p <- p + labs(subtitle = paste("Configuration:", sim_type))
    
    return(p)
  }), storm_names)

  # Return the list of plots for this specific model config
  return(p_list) 
  
}), names(pme_list))

all_storms <- names(all_pme_plots$sim)

# Create a list of side-by-side comparisons
comparison_plots <- setNames(lapply(all_storms, function(storm) {
  
  combo_plot <- all_pme_plots$sim[[storm]] + all_pme_plots$sufi[[storm]] +
    plot_annotation(title = paste("Peak Magnitude Error Comparison:", storm))
  
  return(combo_plot)
}), all_storms)

comparison_plots

```

## Extract worst performing sites

```{r}

pme_dt_list <- lapply(pme_list, function(model_config){
  
  pme_dt <- data.table(
    unsplit_name = names(unlist(model_config)),
    pme = unlist(unlist(model_config))
  )

  pme_dt[, c("Storm", "G2G.ID") := tstrsplit(unsplit_name, "\\.")]
  pme_dt[, unsplit_name := NULL]
  setcolorder(pme_dt, c("Storm", "G2G.ID", "pme"))
  merged_pme <- merge(pme_dt, wales_cdata, by = "G2G.ID")[, .(G2G.ID, NRFA.ID, Site.No.,Site.Name., Storm, pme, CATCHMENTSIZE)]
  merged_pme <- merged_pme[complete.cases(merged_pme[,"pme"])] ## remove NA
  return(merged_pme)
})

worst_pme_sim <- pme_dt_list$sim[, .(
  sim_mean_abs_pme = mean((abs(pme)), na.rm = TRUE),
  sim_mean_pme = mean(pme, na.rm = TRUE)), by = .(G2G.ID, Site.No.,Site.Name., NRFA.ID, CATCHMENTSIZE)]

worst_pme_sufi <- pme_dt_list$sufi[, .(
  sufi_mean_abs_pme = mean((abs(pme)), na.rm = TRUE),
  sufi_mean_pme = mean(pme, na.rm = TRUE)), by = .(G2G.ID, Site.No.,Site.Name., NRFA.ID, CATCHMENTSIZE)]


merged_pme <- merge(
  worst_pme_sim,
  worst_pme_sufi,
  by = c("G2G.ID", "Site.No.", "Site.Name.", "NRFA.ID", "CATCHMENTSIZE")
)

merged_pme <- merged_pme[order(sim_mean_pme,sufi_mean_pme,)]

fwrite(merged_pme,"../data/pme_sim_sufi_events_2018_2021.csv")
 

```

## Explore impcat of ratings

Note here that rating change means that the observed outputs at those sites were using the old rating curve, not the new one. <br>

```{r echo = FALSE}

# catchment_map <- st_read("P:/NEC04218_Ungauged/releases/ffc/releases/G2G_shapefiles/20250610_g2g_ffc_v2_0")
# wales <- st_transform(wales, 27700) # Ensure CRS is BNG (should already be, but safe to enforce)

rating_change_dt <- fread("../data/rating_change_g2g.csv")
pme_dt_sufi <- pme_dt_list$sufi

pme_dt_sufi_rating <- merge(pme_dt_sufi, rating_change_dt, by = "G2G.ID")

pme_dt_sufi_rating[, rating_group := factor(
  new_rating,
  levels = c(0, 1),
  labels = c("No change in rating curve", "Using old rating")
)]

q1 <- quantile(pme_dt_sufi_rating$pme, 0.25, na.rm = TRUE)
q3 <- quantile(pme_dt_sufi_rating$pme, 0.75, na.rm = TRUE)
iqr <- q3 - q1

pme_clean <- pme_dt_sufi_rating[
  pme >= (q1 - 1.5 * iqr) & pme <= (q3 + 1.5 * iqr)
]
counts <- pme_clean[, .(
  n = .N,
  y = max(pme, na.rm = TRUE)
), by = rating_group]

counts[, y := y + 0.05 * diff(range(pme_clean$pme, na.rm = TRUE))]

ggplot(pme_clean, aes(x = rating_group, y = pme, fill = rating_group)) +
  geom_boxplot(alpha = 0.7) +
  geom_text(
    data = counts,
    aes(x = rating_group, y = y, label = paste0("n = ", n)),
    inherit.aes = FALSE,
    vjust = 0
  ) +
  labs(
    x = "Rating change",
    y = "PME",
    title = "PME by whether site has changed rating since 2022 (SIM)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

pme_dt_sufi <- pme_dt_list$sufi

pme_dt_sufi_rating <- merge(pme_dt_sufi, rating_change_dt, by = "G2G.ID")

pme_dt_sufi_rating[, rating_group := factor(
  new_rating,
  levels = c(0, 1),
  labels = c("No change in rating curve", "Using old rating")
)]

q1 <- quantile(pme_dt_sufi_rating$pme, 0.25, na.rm = TRUE)
q3 <- quantile(pme_dt_sufi_rating$pme, 0.75, na.rm = TRUE)
iqr <- q3 - q1

pme_clean <- pme_dt_sufi_rating[
  pme >= (q1 - 1.5 * iqr) & pme <= (q3 + 1.5 * iqr)
]
counts <- pme_clean[, .(
  n = .N,
  y = max(pme, na.rm = TRUE)
), by = rating_group]

counts[, y := y + 0.05 * diff(range(pme_clean$pme, na.rm = TRUE))]

ggplot(pme_clean, aes(x = rating_group, y = pme, fill = rating_group)) +
  geom_boxplot(alpha = 0.7) +
  # geom_text(
  #   data = counts,
  #   aes(x = rating_group, y = y, label = paste0("n = ", n)),
  #   inherit.aes = FALSE,
  #   vjust = 0
  # ) +
  labs(
    x = "Rating change",
    y = "PME",
    title = "PME by whether site has changed rating since 2022 (SUFI)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```


## Test impact of slope heterogeneity


```{r}

wales_dem <- terra::rast("../data/rasters/wales_dem_50m.tif")

slope_rast <- terra::terrain(wales_dem, v = "slope", unit = "degrees")
tri_rast   <- terra::terrain(wales_dem, v = "TRI")

catchments <- st_read("../data/20250610_g2g_ffc_v2_0/fullcats_g2g_ffc_v2_0_loc.shp")

catchments_v <- terra::vect(catchments)

print("Extracting metrics for each catchment...")
slope_extracted <- terra::extract(slope_rast, catchments_v, fun = mean, na.rm = TRUE)
tri_extracted   <- terra::extract(tri_rast, catchments_v, fun = mean, na.rm = TRUE)
tri_range_extracted   <- terra::extract(tri_rast, catchments_v, fun = sd, na.rm = TRUE) ## range might be more interesting for heterogeneity ?

terrain_dt <- data.table(
  G2G.ID     = catchments$g2g_id, 
  Mean_Slope = slope_extracted[[2]],  # The second column holds the actual raster values
  Mean_TRI   = tri_extracted[[2]],
  sd_TRI = tri_range_extracted[[2]]
)

wales_cdata <- merge(wales_cdata, terrain_dt, by = "G2G.ID", all.x = TRUE)

# head(wales_cdata[, .(G2G.ID, River.Name, Mean_Slope, Mean_TRI)])
# merged_pme <- merge(pme_dt_list$sim, pme_dt_list$sufi)

# pme_tri_list <- lapply(pme_dt_list, function(x) merge(x, terrain_dt, by = "G2G.ID"))



# plot(log(abs(sufi_mean_pme)) ~ sd_TRI, data = pme_tri)


```


## River gauge metrics

```{r}

river_gauges <- fread("../data/rg_info_2022.csv")

gauges_sf <- st_as_sf(
  river_gauges, 
  coords = c("X", "Y"), 
  crs = 27700
)

gauges_v <- terra::vect(gauges_sf)

print("Generating proximity-to-gauge distance map...")
gauge_dist_rast <- terra::distance(wales_dem, gauges_v)

print("Extracting mean gauge distance per catchment...")
dist_extracted <- terra::extract(gauge_dist_rast, catchments_v, fun = mean, na.rm = TRUE)
gauge_dist_dt <- data.table(
  G2G.ID     = catchments$g2g_id, 
  mean_dist = dist_extracted[[2]],  
  mean_dist_km = dist_extracted[[2]] /1000 
)


# # pme_dist_tri <- merge(gauge_dist_dt, pme_tri, by = "G2G.ID")
# pme_dist_tri <- lapply(pme_tri_list, function(x) merge(x, gauge_dist_dt, by = "G2G.ID"))


# lapply(pme_dist_tri, function(x){
#   x[, log_abs_pme := log(abs(pme))]
#   x[, Mean_Gauge_Dist_Km := mean_dist / 1000]
# })


# plot(log(abs(sufi_mean_pme)) ~ mean_dist, data = pme_dist)

wales_cdata <- merge(wales_cdata, gauge_dist_dt, by = "G2G.ID", all.x = TRUE)

```


## Incorporating rainfall data

- Total event rainfall depth (sum over the event window)
- Peak rainfall intensity (max value, or max over a rolling window e.g. 1hr/6hr — intensity matters more than total depth for flashy response)
- Time-to-peak rainfall (for comparing against time-to-peak flow — see timing below)
- Rainfall duration above some threshold

```{r}

rf_dat <- fread("../data/catchment_average_rg_precip.csv")

storm_names <- names(events_list[[1]])
rf_event_dat <- list()
for (storm in storm_names){
  event_period <- c(min(events_list$sim[[storm]]$mod$DATE_TIME), max(events_list$sim[[storm]]$mod$DATE_TIME))
  ids_to_keep <- gsub("_Mod", "", names(events_list$sim[[storm]]$mod))[-1] ## remove DATE_TIME column
  rf_event_dat[[storm]] <- rf_dat[date >= event_period[1] & date <= event_period[2],  c("date",ids_to_keep), with = FALSE]
  setnames(rf_event_dat[[storm]], "date", "DATE_TIME") ## set it to same name as g2g output
}

rf_event_sum <- lapply(rf_event_dat, function(x) apply(x[, !"DATE_TIME"], 2, sum))
rf_event_sum_dt_list <- lapply(rf_event_sum, function(x) data.table(G2G.ID = names(x), precip = as.numeric(x)))

rf_event_sum_dt <- rbindlist(rf_event_sum_dt_list, idcol = "Storm")


```

## incorporating urban extent

- weight urban as 1, suburban as 0.5.


```{r}

suburban_lcm <- rast("../data/rasters/suburban_LCM2019.asc")
urban_lcm <- rast("../data/rasters/urban_LCM2019.asc")

wales_catchments <- st_read("../data/wales_catchments_sf")
catchments_v <- terra::vect(wales_catchments)

suburb_extracted <- terra::extract(suburban_lcm, catchments_v, fun = mean, na.rm = TRUE)
urb_extracted <- terra::extract(urban_lcm, catchments_v, fun = mean, na.rm = TRUE)

urb_sub_combined <- data.table(G2G.ID = catchments_v$g2g_id, urban.suburban.cover = (urb_extracted[[2]]* 1) + (suburb_extracted[[2]]* 0.5) )
wales_cdata <- merge(wales_cdata, urb_sub_combined, by = "G2G.ID")

# urban_dt <- data.table(
#   G2G.ID     = catchments$g2g_id, 
#   Mean_Slope = slope_extracted[[2]],  # The second column holds the actual raster values
#   Mean_TRI   = tri_extracted[[2]],
#   sd_TRI = tri_range_extracted[[2]]
# )

```


## NRW Threshold exceedance

```{r}
max_discharge_sim_sufi <- lapply(events_list, lapply, extract.peak.discharge)

nrw_threshold <- fread("../data/nrw_alert_warning_flow_thresh.csv", fill =TRUE, sep=",")


flood_alert <- setNames(
  nrw_threshold[!is.na(`Flood Alert`), `Flood Alert`],
  nrw_threshold[!is.na(`Flood Alert`), G2G.ID]
)

flood_warning <- setNames(
  nrw_threshold[!is.na(`Flood Warning`), `Flood Warning`],
  nrw_threshold[!is.na(`Flood Warning`), G2G.ID]
)

severe_flood_warning <- setNames(
  nrw_threshold[!is.na(`Severe Flood Warning`), `Severe Flood Warning`],
  nrw_threshold[!is.na(`Severe Flood Warning`), G2G.ID]
)

flood_thresh <- list(alert = flood_alert, warning=flood_warning, severe_warning = severe_flood_warning)


max_discharge_sim_sufi_filt <- lapply(flood_thresh, function(level){
  lapply(max_discharge_sim_sufi, lapply, function(x){
    g2g_sites <- names(level)
    mod <- x$mod
    obs <- x$obs
    mod_sites <- gsub("_Mod", "", names(mod))
    obs_sites <- gsub("_Obs", "", names(obs))
    filt <- mod_sites %in% g2g_sites
    mod_filt <- mod[filt]
    obs_filt <- obs[filt] ## filter so theyre matching
    return(list(obs = obs_filt, mod = mod_filt))
  })
})

```





#### With tolerance

```{r}

# max_discharge_sim_sufi <- lapply(events_list, lapply, extract.peak.discharge)

nrw_thresh_exceed <- Map(function(max_disch, thresh_type){
  lapply(max_disch, lapply, pod.far.tol, threshold = thresh_type)
}, max_discharge_sim_sufi_filt, flood_thresh)

nrw_thresh_exceed <- lapply(nrw_thresh_exceed, lapply,lapply, merge, y = wales_cdata, by = "G2G.ID")
nrw_thresh_exceed_dt <- lapply(nrw_thresh_exceed, lapply, rbindlist, idcol = "Storm")
nrw_thresh_exceed_dt_rg <- lapply(nrw_thresh_exceed_dt, lapply, merge, rf_event_sum_dt, by = c("G2G.ID", "Storm")) ## add rain gauge data


storm_names <- unique(nrw_thresh_exceed_dt$sim$Storm)

detection_data <- lapply(nrw_thresh_exceed_dt_rg$alert, function(dt){
  dt[outcome %in% c("hit", "near miss", "miss")]
})



detection_data <- lapply(detection_data, function(dt){
  dt[, `:=`(
  hit_strict  = as.integer(outcome == "hit"),                    # near miss counts as a miss
  hit_lenient = as.integer(outcome %in% c("hit", "near miss"))   # near miss counts as a hit
)]
})


detection_data <- lapply(detection_data, function(x){
  x[, urban_sc := scale(urban.suburban.cover)]
  x[, precip_sc := scale(precip)]
  x[, catchmentsize_sc := scale(CATCHMENTSIZE)]
  x[,terrian_sc := scale(sd_TRI)]
})

detection_strict <- glmer(hit_strict ~ precip_sc + catchmentsize_sc +  urban_sc + terrian_sc +(1 | G2G.ID) + (1 | Storm),
                       data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

detection_lenient <- glmer(hit_lenient ~ precip_sc + catchmentsize_sc +  urban_sc  +terrian_sc+(1 | G2G.ID) + (1 | Storm),
                        data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

exp(fixef(detection_strict)); exp(fixef(detection_lenient))

summary(detection_strict)

summary(detection_lenient)

detection_interaction <- glmer(hit_strict ~ precip_sc * urban_sc+ catchmentsize_sc + terrian_sc +(1 | G2G.ID) + (1 | Storm),
                       data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

summary(detection_interaction)

# anova(detection_strict, detection_interaction)
```

Detection probability increased significantly with precipitation, catchment size, and urban/suburban cover. Inclusion of a precipitation × urbanisation interaction did not improve model performance (LRT: χ² = 0.56, df = 1, p = 0.45; ΔAIC = +1.43), suggesting that the influence of precipitation on detection probability was not dependent on urbanisation.



## plots of detection vs predictors

```{r echo = FALSE}

library(ggplot2)
library(data.table)


# plot_dt[, log_catchment := log(CATCHMENTSIZE)]

p <- ggplot(nrw_thresh_exceed_dt_rg$alert$sim, aes(x = CATCHMENTSIZE, y = outcome, color = outcome)) +
  geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
  labs(
    x = "Catchment size km^2",
    y = "Detection category",
    title = "Detection categories by catchment size (SIM)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")


q <- ggplot(nrw_thresh_exceed_dt_rg$alert$sufi, aes(x = CATCHMENTSIZE, y = outcome, color = outcome)) +
  geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
  labs(
    x = "Catchment size km^2",
    y = "Detection category",
    title = "Detection categories by catchment size (SUFI)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")


p + q
p <- ggplot(nrw_thresh_exceed_dt_rg$alert$sim, aes(x = precip, y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "summed average gauge interpolated precipitiation",
  y = "Detection category",
  title = "Detection categories by precipitiation (Sim)"
) +
theme_minimal() +
theme(legend.position = "none")


q <- ggplot(nrw_thresh_exceed_dt_rg$alert$sufi, aes(x = precip, y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "summed average gauge interpolated precipitiation",
  y = "Detection category",
  title = "Detection categories by precipitiation (SUFI)"
) +
theme_minimal() +
theme(legend.position = "none")

p + q

p <- ggplot(nrw_thresh_exceed_dt_rg$alert$sim, aes(x = urban.suburban.cover, y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "Urban/suburban land coverage",
  y = "Detection category",
  title = "Detection categories by urban/suburban cover (Sim)"
) +
theme_minimal() +
theme(legend.position = "none")


q <- ggplot(nrw_thresh_exceed_dt_rg$alert$sufi, aes(x = urban.suburban.cover, y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "Urban/suburban land coverage",
  y = "Detection category",
  title = "Detection categories by urban/suburban cover (SUFI)"
) +
theme_minimal() +
theme(legend.position = "none")


p+q

```


# Peak timing threshold

```{r}




```


# QMED Threshold modelling

## extract 


```{r}

max_discharge_sim_sufi <- lapply(events_list, lapply, extract.peak.discharge)
qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grid_list <- lapply(qt_grid_paths, read.ascii)
names(qt_grid_list) <- qt_val ## name by qt
qt_dt <- make.qt.csv(qt_grid_list, cdata = wales_cdata)

qmed <-qt_dt[, .(G2G.ID, qmed)]$qmed
names(qmed) <- qt_dt$G2G.ID
qmed_pod_far_tol <- lapply(max_discharge_sim_sufi, lapply, pod.far.tol, threshold = qmed)


qmed_pod_far_tol <- lapply(qmed_pod_far_tol, lapply, merge, y = wales_cdata, by = "G2G.ID")
qmed_pod_far_tol_dt <- lapply(qmed_pod_far_tol, rbindlist, idcol = "Storm")
qmed_pod_far_tol_dt_rg <- lapply(qmed_pod_far_tol_dt, merge, rf_event_sum_dt, by = c("G2G.ID", "Storm")) ## add rain gauge data


storm_names <- unique(qmed_pod_far_tol_dt$sim$Storm)

comparison_plots <- setNames(lapply(storm_names, function(storm_name) {
  p_sim  <- plot.wales.nrw.thresh.tol(qmed_pod_far_tol_dt$sim,  wales, storm.name = storm_name) +
    labs(subtitle = "SIM")

  p_sufi <- plot.wales.nrw.thresh.tol(qmed_pod_far_tol_dt$sufi, wales, storm.name = storm_name) +
    labs(subtitle = "SUFI")

  p_sim + p_sufi +
    plot_annotation(title = paste("Storm:", storm_name))
}), storm_names)
comparison_plots

```

```{r eval = FALSE}

create.csi.plot <- function(dt, lenient = FALSE){
  site_counts <- dt[, .N, by = .(G2G.ID, outcome)]
  site_counts_wide <- dcast(site_counts, G2G.ID ~ outcome, value.var = "N", fill = 0)


  site_counts_wide[, CSI_strict := hit / (hit + miss + `false alarm`)]
  site_counts_wide[, CSI_lenient :=
    (hit + `near miss`) / (hit + `near miss` + miss + `false alarm`)]
    required_cols <- c("hit", "near miss", "miss", "false alarm")

  for (col in required_cols) {
    if (!col %in% names(site_counts_wide)) {
      site_counts_wide[, (col) := 0]
    }
  }

  if (!lenient){
      csi <- site_counts_wide[, .(G2G.ID, CSI_strict)]
  } else {
    csi <- site_counts_wide[, .(G2G.ID, CSI_lenient)]
  }

  p <- plot.single.point(
  site_counts_wide$G2G.ID,
  wales = wales,
  cdata = wales_cdata,
  fill.variable = csi
) +
  labs(
    fill = "CSI",
    size = expression("Catchment size (" * km^2 * ")")
  ) +
  theme(
    # transparent background
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),

    # remove grid
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),

    # remove axes
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),

    # optional: remove title
    plot.title = element_blank()
  )
  return(p)
}

sufi_csi_qmed <- create.csi.plot(qmed_pod_far_tol_dt$sufi, lenient = FALSE)


sim_csi_qmed <- create.csi.plot(qmed_pod_far_tol_dt$sim, lenient = FALSE)

# ggsave(
#           filename = "../data/figures/sim_csi_qmed.png",
#           plot = sim_csi_qmed,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300,
#           bg = "transparent"
#         )
# ggsave(
#           filename = "../data/figures/sufi_csi_qmed.png",
#           plot = sufi_csi_qmed,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300,
#           bg = "transparent"
#         )


sufi_csi_qmed_len <- create.csi.plot(qmed_pod_far_tol_dt$sufi, lenient = TRUE)


sim_csi_qmed_len <- create.csi.plot(qmed_pod_far_tol_dt$sim, lenient = TRUE)

# ggsave(
#           filename = "../data/figures/sim_csi_qmed_lenient.png",
#           plot = sim_csi_qmed_len,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300,
#           bg = "transparent"
#         )
# ggsave(
#           filename = "../data/figures/sufi_csi_qmed_lenient.png",
#           plot = sufi_csi_qmed_len,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300,
#           bg = "transparent"
#         )
```


## Model QMED detection

```{r}


detection_data <- lapply(qmed_pod_far_tol_dt_rg, function(dt){
  dt[outcome %in% c("hit", "near miss", "miss")]
})



detection_data <- lapply(detection_data, function(dt){
  dt[, `:=`(
  hit_strict  = as.integer(outcome == "hit"),                    # near miss counts as a miss
  hit_lenient = as.integer(outcome %in% c("hit", "near miss"))   # near miss counts as a hit
)]
})


detection_data <- lapply(detection_data, function(x){
  x[, urban_sc := scale(urban.suburban.cover)]
  x[, precip_sc := scale(precip)]
  x[, catchmentsize_sc := scale(CATCHMENTSIZE)]
  x[,terrian_sc := scale(sd_TRI)]
  x[,rg_dist_sc := scale(mean_dist_km)]
})

detection_strict <- glmer(hit_strict ~ precip_sc + catchmentsize_sc +  urban_sc + rg_dist_sc + terrian_sc+(1 | G2G.ID) + (1 | Storm),
                       data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))


detection_lenient <- glmer(hit_lenient ~ precip_sc + catchmentsize_sc +  urban_sc  +(1 | G2G.ID) + (1 | Storm),
                        data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

summary(detection_strict)
summary(detection_lenient)


detection_interaction <- glmer(hit_strict ~ precip_sc * urban_sc+ catchmentsize_sc + terrian_sc +(1 | G2G.ID) + (1 | Storm),
                       data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

summary(detection_interaction)

# anova(detection_strict, detection_interaction)

```


## Plots qmed with predicors

```{r}


p <- ggplot(qmed_pod_far_tol_dt_rg$sim, aes(x = scale(precip), y = outcome, color = outcome)) +
  geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
  labs(
    x = "Scaled Catchment size km^2",
    y = "Detection category",
    title = "Detection categories by catchment size (SIM)"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
      plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA))



q1 <- ggplot(qmed_pod_far_tol_dt_rg$sufi, aes(x = scale(CATCHMENTSIZE), y = outcome, color = outcome)) +
  geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
  labs(
    x = "Scaled Catchment size km^2",
    y = "Detection category",
    title = "Detection categories by catchment size (SUFI)"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
  plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA))






p + q
p <- ggplot(qmed_pod_far_tol_dt_rg$sim, aes(x = precip, y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "summed average gauge interpolated precipitiation",
  y = "Detection category",
  title = "Detection categories by precipitiation (SUFI)"
) +
theme_minimal() +
theme(legend.position = "none")


q2 <- ggplot(qmed_pod_far_tol_dt_rg$sufi, aes(x = scale(precip), y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "Scaled summed average gauge \n interpolated precipitiation",
  y = "Detection category",
  title = "Detection categories by precipitiation (SUFI)"
) +
theme_minimal() +
theme(legend.position = "none",
plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA))

p + q

p <- ggplot(qmed_pod_far_tol_dt_rg$sim, aes(x = urban.suburban.cover, y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "summed average gauge interpolated precipitiation",
  y = "Detection category",
  title = "Detection categories by urban/suburban cover (SUFI)"
) +
theme_minimal() +
theme(legend.position = "none")


q3 <- ggplot(qmed_pod_far_tol_dt_rg$sufi, aes(x = scale(urban.suburban.cover), y = outcome, color = outcome)) +
geom_jitter(height = 0.15, width = 0, alpha = 0.6, size = 2) +
labs(
  x = "Scaled urban/suburban score",
  y = "Detection category",
  title = "Detection categories by urban/suburban cover (SUFI)"
) +
theme_minimal() +
theme(legend.position = "none",
plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA))


p+q

ggsave(
          filename = "../data/figures/sufi_boxplot_hit.png",
          plot = q1 +q2 +q3,
          width = 15,
          height = 8,
          units = "in",
          dpi = 300,
          bg = "transparent"
        )


```


# things to do 14/07

- see if there are any catchments that perform badly with nrw threshold, pme, and kge
where are they on the map?


# things to do

how are sites performing in general. longer period hydrograph covering few years. KGE statistics at these sites. is that site performing well, or is it the high flows that we're getting.
3 years, across 2018.
Look at NRFA, see if there is anything particularly noticable.


- are there reservoirs
- gauges upstream feeding into it.
- headwater any 
- can we map slopes (tomography)


Also the rainfall that is going in, needs checking.

Raingauges.
- one idea is the amount of raingauges in a certain radius, were g2g site is the centroid.
-  Catchment size split up; altert warnings per catchment
Elwy, putting in thresholds
Gridded outputs.
  - john making gridded outputs
  - can look at different high flows at different parts of the river.
Some of the other storm events, pre 2018.
Do analyses per catchment size.
Pull Ma from midland.
