---
title: "QT Exceedannce plots  for Storm events (Wales 2018-2021)"
subtitle: "Sim & SUFI data"
author: "Caleb Scutt"
format:
  html:
    page-layout: full
    toc: true
    code-fold: true
    fig-width: 10
    fig-height: 5
    embed-resources: true
    theme: flatly
  typst:
    toc: true
    fig-width: 7
    fig-height: 5
    margin:
      x: 1in
      y: 1in
    fontsize: 11pt
    section-numbering: 1.1.1
    keep-typ: false
execute:
  warning: false
  message: false
---


```{r echo = FALSE}
## load packages and functions
library(data.table)
library(ggplot2)
library(lubridate)
library(gtools)
library(sf)
library(grid)
library(gridExtra)
library(patchwork)
library(gt)
library(terra)
library(lme4)
source("../functions/load.g2g.runs.R")
source("../functions/extract.g2g.grid.R")
# source("../functions/qt.exceed.R")
source("../functions/plot.wales.R")
source("../functions/obs.mod.stats.R")
```


<!-- ## Read in data

* `mod_ts` is the g2g modelled flow data at each g2g site
* `obs_ts` is the observed flow data at each g2g site
* `events` are the notable storm events in Wales (`events_pre_2022` are events we have data for)
* `catchments` are the metadata for each catchment (region, area, river name, easting/northing, g2g.id). `wales_cdata` is this filtered by just Welsh catchments
* The following sites are missing from g2g model output, but present in catchment data: 055025_TG_325_Mod, 57809_Mod 60005_Mod, 066004_TG_9512_Mod, 067005_TG_116_Mod, 067023_TG_117_Mod, 055025_TG_325_Mod, 60005_Mod. -->

```{r, include = FALSE}
g2g_output_2018_2022 <- load.g2g.data(dates = c(2018, 2021), sufi = TRUE)
obs_ts <- g2g_output_2018_2022$obs
mod_ts <- g2g_output_2018_2022$mod
## read in the events
events <- fread("../data/notable_events.csv")
events_pre_2022 <- events[Year <= 2021,]
## read in the catchment metadata
catchments <- fread("../data/cdata/catchment_cdata_EA-NRW.csv", fill = Inf)
## filter by wales
wales <- st_read("../data/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")
wales <- st_transform(wales, 27700)

catchments_sf <- st_as_sf(
  catchments,
  coords = c("WISKI.EASTING", "WISKI.NORTHING"),
  crs = 27700,
  remove = FALSE
)
wales_poly <- st_union(wales)
sf_g2g_ids <- unique(c(catchments_sf[st_intersects(catchments_sf, wales_poly, sparse = FALSE), ]$G2G.ID, catchments[Region. == "Wales"]$G2G.ID)) ## use both ids coded as Wales, plus anything inside the polygon
wales_cdata <- catchments[G2G.ID %in% sf_g2g_ids]## filter by wales


```


<!-- Here we make a nested list, with top level of list representing each storm event from `events`. <br>
We subset Wales into North, Mid and South by these Northing values (can be adjusted). These were used because some sites were not categorised by Area (e.g South East [CY]) <br>
We subset the time series 7 days either side of the event beginning/end. <br> -->


```{r message = FALSE, warning = FALSE, include = FALSE, echo = TRUE}

events_list <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  NULL, exclude.non.notable.sites = FALSE)

```


<!-- ## Read in QT data

QT data is extracted using functions from `functions/griddedData_class_jrw.py` converted to R.
These functions can be re-used when more catchment data is added.
Note that these QT values are calculated pre-2022. -->



```{r, message = FALSE, include = FALSE}
## this chunk extracts the g2g qt vals
qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grid_list <- lapply(qt_grid_paths, read.ascii)
names(qt_grid_list) <- qt_val ## name by qt
qt_dt <- make.qt.csv(qt_grid_list, cdata = wales_cdata)
```


## Extract the peak discharge at each flood event

* Peak discharge extracted at each event at each g2g site for observed (what actually happened) and modelled (sim and sufi).

```{r, include = FALSE}

sim_max_discharge_qt <- list()
for (t_val in qt_val) {
  sim_max_discharge_qt[[t_val]] <- lapply(events_list$sim, function(event_mod_obs) {
    extract.peak.discharge(event_mod_obs, qt_dt, t_val)
  })
}

sufi_max_discharge_qt <- list()
for (t_val in qt_val) {
  sufi_max_discharge_qt[[t_val]] <- lapply(events_list$sufi, function(event_mod_obs) {
    extract.peak.discharge(event_mod_obs, qt_dt, t_val)
  })
}

max_discharge_qt <- list(sim = sim_max_discharge_qt, sufi = sufi_max_discharge_qt)

qt_exceeded <- lapply(max_discharge_qt,lapply, lapply, thresh.exceed)


qt_exceeded_dt_list <- lapply(qt_exceeded, lapply, function(qt){
    dt_exceed <- data.table(
    unsplit_name     = names(unlist(qt)),
    Exceeded = unlist(qt)
    )
    dt_exceed[, c("Storm", "Data_Type", "G2G.ID") := tstrsplit(unsplit_name, "\\.", keep = 1:3)]
    dt_exceed[, G2G.ID := gsub("_Obs|_Mod", "", G2G.ID)]
    dt_exceed[, unsplit_name := NULL]
    setcolorder(dt_exceed, c("Storm", "Data_Type", "G2G.ID", "Exceeded"))
    return(dt_exceed)
})

merged_dat <- lapply(qt_exceeded_dt_list, lapply, merge, qt_dt, by = "G2G.ID")

```


```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
# wales <- st_read("W:/hymod/Hydro-JULES/HJ Internships/2026 - NRW G2G/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")# Read Wales constituency polygons
# 

q_levels <- c("q1", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")

highest_qt_exceeded_list <- lapply(merged_dat, function(model_config){
    exceed_dt <- rbindlist(model_config, idcol = "Threshold")

    exceed_dt[, Threshold := gsub("qmed", "q1", Threshold)] ## sub qmed for q1 so it can be ordered correctly in next step

    exceed_dt[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

    exceed_dt[, keep_site_storm := any(Exceeded == TRUE, na.rm = TRUE), by = .(Storm, G2G.ID)] ## identify sites exceeded in any of 3 data (obs, sim sufi)
    filtered_dt <- exceed_dt[keep_site_storm == TRUE] ## filter these sites

    highest_qt_exceeded <- filtered_dt[, .(
      Threshold = if (any(Exceeded == TRUE, na.rm = TRUE)) {
        as.character(max(Threshold[Exceeded == TRUE], na.rm = TRUE))
      } else {
        "None"
      },
      Site         = first(Site),
      G2G_Easting  = first(G2G_Easting),
      G2G_Northing = first(G2G_Northing)
    ), by = .(Storm, G2G.ID, Data_Type)]

    highest_qt_exceeded[, Threshold := gsub("^q1$", "qmed", Threshold)]

    final_levels <- c("None", "qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")
    highest_qt_exceeded[, Threshold := factor(Threshold, levels = final_levels, ordered = TRUE)]

    # Optional: Sort it cleanly for viewing
    setorder(highest_qt_exceeded, Storm, G2G.ID, Data_Type)
    return(highest_qt_exceeded)
})

q_levels <- c("qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000") ## reassign q_levels back to normal


```

## QT Exceeded plots

* The following plots show the highest flow threshold exceeded at each site, for each storm.

```{r echo = FALSE}
#| fig-width: 14
#| fig-height: 5


for (name in unique(highest_qt_exceeded_list[[1]]$Storm)) {
  p <- plot.wales.exceed(highest_qt_exceeded_list$sufi, wales, storm.name = name, "obs")
  q <- plot.wales.exceed(highest_qt_exceeded_list$sim, wales, storm.name = name, "sim")
  r <- plot.wales.exceed(highest_qt_exceeded_list$sufi, wales, storm.name = name, "sufi")

  pqr <- list(obs = p, sim = q,  sufi = r)

  # lapply(names(pqr), function(config){

  #   plot <- pqr[[config]]

  #   ggsave(
  #         filename = file.path(paste0("../data/figures/qt_exceed_maps"), paste0(name, "_", config, "_qt_exceed_map", ".png")),
  #         plot = plot,
  #         width = 8,
  #         height = 6,
  #         units = "in",
  #         dpi = 300
  #       )
  # })

  print(
    (p + q + r) +
      plot_annotation(title = paste("Storm:", name))
  )
}

```

## Notable catchments hydrographs

These are hydrographs taken from the notable catchments highlighted by NRW for each storm event. <br>

```{r echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
#| fig-width: 15
#| fig-height: 10
source("../functions/make.hydrograph.r")
events_list_notable_catchments <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  NULL, exclude.non.notable.sites = TRUE)

```

```{r echo = FALSE}
#| fig-align: left
#| fig-height: 6

library(patchwork)


events_list_notable_catchments_pivoted <- list()
for (event in names(events_list_notable_catchments[[1]])) {
  sim <- events_list_notable_catchments$sim[[event]]$mod
  sufi <- events_list_notable_catchments$sufi[[event]]$mod
  obs <- events_list_notable_catchments$sufi[[event]]$obs
  events_list_notable_catchments_pivoted[[event]] <- list(sim = sim, sufi = sufi, obs = obs)
}


```


```{r echo=FALSE, results='asis'}

for (event in names(events_list_notable_catchments_pivoted)) {

  event_data <- events_list_notable_catchments_pivoted[[event]]

  # Print the main event section header (HTML heading prevents bracket bugs)
  cat(paste0("\n\n<h3>Storm Event: ", event, "</h3>\n\n"))

  # Extract G2G IDs
  g2g_ids <- gsub("_Mod", "", colnames(event_data$sim)[2:length(colnames(event_data$sim))])

  # Ensure CATCHMENTSIZE is numeric
  ord_catchment_size_river <- wales_cdata[G2G.ID %in% g2g_ids]
  ord_catchment_size_river[, CATCHMENTSIZE := as.numeric(CATCHMENTSIZE)]

  # Find files
  ev_files <- data.table(
    path = list.files(
      path = file.path("../data/figures/wales_events_2018_2021_hydrographs"),
      pattern = "\\.png$",
      full.names = TRUE
    )
  )

  if (nrow(ev_files) == 0) next

  ev_files[, name := tools::file_path_sans_ext(basename(path))]
  ev_files[, c("event_file", "g2g_id") := tstrsplit(name, "__", fixed = TRUE)]
  ev_files[, g2g_id := as.character(g2g_id)]

  ev_files <- ev_files[event_file == event]
  if (nrow(ev_files) == 0) next

  ev_files <- merge(
    ev_files,
    ord_catchment_size_river[, .(G2G.ID, River.Name, CATCHMENTSIZE)],
    by.x = "g2g_id",
    by.y = "G2G.ID",
    all = FALSE
  )

  setorder(ev_files, River.Name, CATCHMENTSIZE)

  # Generate HTML sub-headers and inject side-by-side images
  for (rv in unique(ev_files$River.Name)) {
    
    # Using an HTML <h4> tag instead of #### protects the brackets in TAFF[SE](WA)
    cat(paste0("\n\n<h4>River System: ", rv, "</h4>\n\n"))
    
    paths <- ev_files[River.Name == rv, path]
    
    # Wrap images in a simple div container, and set each image to 49% width 
    # so they naturally sit side-by-side
    cat("<div style='width: 100%;'>\n")
    for (img_path in paths) {
      cat(paste0("<img src='", img_path, "' style='width: 49%; display: inline-block; padding: 2px;'>\n"))
    }
    cat("</div>\n\n")
  }
}
```

<!-- ## Hydrographs (obs vs modelled comparison)

### Largest flow mismatches in model vs obs

Hydrographs shown for where there is the largest discrepancies in the QT exceeded (`worst_mismatches$rank_diff>=4`).
Note that the storm events are lumped together here. <br>
Top 10 plotted -->

```{r echo = FALSE}
#| fig-width: 15
#| fig-height: 20


comp_dt_list <- lapply(highest_qt_exceeded_list, function(x) {
  comp_dt <- dcast(x, Storm + G2G.ID + Site + G2G_Easting + G2G_Northing ~ Data_Type, value.var = "Threshold" )
  comp_dt[, rank_mod := as.numeric(mod)] ## ranked where qmed is 1, q1000 is 10
  comp_dt[, rank_obs := as.numeric(obs)]
  comp_dt[is.na(rank_mod), rank_mod := 0]
  comp_dt[is.na(rank_obs), rank_obs := 0]
  comp_dt[, rank_diff := abs(rank_mod - rank_obs)] ## calculate rank differences
})


```

```{r eval = FALSE}

worst_mismatches <- comp_dt[!is.na(rank_diff)][order(-rank_diff)]
worst_mismatches <-worst_mismatches[worst_mismatches$rank_diff >=4] ## filter by where qt was a difference of 3


mismatch_plots <- list()
n_mismatch <- min(10, nrow(worst_mismatches))
for (i in 1:n_mismatch) {
  event <- worst_mismatches[i, Storm]
  g2g_id <- worst_mismatches[i, G2G.ID]
  event_data <- events_list[[event]]
  mismatch_plots[[i]] <- tryCatch(make.hydrograph(event_data, g2g_id, storm.name = event, cdata = wales_cdata, qt.data = filtered_qt), error = function(e) NA)
}
mismatch_plots <- mismatch_plots[!sapply(mismatch_plots, is.na)]

wrap_plots(mismatch_plots, ncol = 2) +
  plot_annotation(title = "Largest flow mismatches (rank_diff >= 4)")


```


<!-- ### False alarms (model exceeds qmed, observed under qmed ) -->

```{r echo = FALSE, warning = FALSE, eval = FALSE}
#| fig-width: 15
#| fig-height: 20

# model false alarm for flood
model_false_positives <- comp_dt[is.na(obs)]
model_false_positives <- model_false_positives[order(-mod)]

false_positive_plots <- list()
for (i in 1:16){
  event <- model_false_positives[i, Storm]
  g2g_id <- model_false_positives[i, G2G.ID]
  event_data <- events_list[[event]]
  false_positive_plots[[i]] <- tryCatch(make.hydrograph(event_data, g2g_id, storm.name = event, cdata = wales_cdata, qt.data = filtered_qt), error = function(e) NA)
} ## some of these are NA because lack of obs data??

false_positive_plots <- false_positive_plots[!sapply(false_positive_plots, is.na)]

grid.arrange(grobs = false_positive_plots)

```

<!-- ### Model misses (observed exceeds qmed, model under qmed ) -->

```{r eval = FALSE}
#| fig-width: 15
#| fig-height: 20

model_miss <- comp_dt[is.na(mod)]
model_miss <- model_miss[order(-obs)]
model_miss_plots <- list()

for (i in 1:10){
  event <- model_miss[i, Storm]
  g2g_id <- model_miss[i, G2G.ID]
  event_data <- events_list[[event]]
  model_miss_plots[[i]] <- make.hydrograph(event_data, g2g_id, storm.name = event, cdata = wales_cdata, qt.data = filtered_qt)
}
grid.arrange(grobs = model_miss_plots)

```


## Prediction accuracy metrics

* POD = probability of detection (closer to 1 = better)
* FAR = false alarm ratio (closer to 0 = better)
* CSI = critical success ratio  (closer to 1 = better)

```{r echo = TRUE}
#| fig-height: 5


# 1. Count the occurrences of each obs/mod combination
sim_heatmap_dt <- comp_dt_list$sim[obs != "None" & mod != "None", .N, by = .(obs, mod)]
sim_heatmap_dt[, `:=`(
  obs = droplevels(obs), ## this removed the unused "None " factor from the data.table
  mod = droplevels(mod)
)] ## 

# 2. Plot the heatmap
p <- ggplot(sim_heatmap_dt, aes(x = obs, y = mod, fill = N)) +
  geom_tile(color = "white", size = 0.5) + 
  geom_text(aes(label = N, color = N > max(N)/2), show.legend = FALSE, size = 4) +
  scale_color_manual(values = c("black", "white")) +
  
  # A clean gradient (white to steelblue works well for hydrology reports)
  scale_fill_distiller(palette = "Blues", direction = 1, name = "Count") +  # Ensure all factor levels display on the axes, even if the count is 0
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  
  labs(
    title = "Sim model vs Observed Discharge Thresholds",
    x = "Observed Threshold (obs)",
    y = "Modelled Threshold (SIM)"
  ) +
  
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Removes default grid lines for a cleaner look
    axis.text.x = element_text(angle = 45, hjust = 1) # Angles x-axis labels if they get crowded
)


sufi_heatmap_dt <- comp_dt_list$sufi[obs != "None" & mod != "None", .N, by = .(obs, mod)]
sufi_heatmap_dt[, `:=`(
  obs = droplevels(obs), ## this removed the unused "None " factor from the data.table
  mod = droplevels(mod)
)] ## 

# 2. Plot the heatmap
q <- ggplot(sufi_heatmap_dt, aes(x = obs, y = mod, fill = N)) +
  geom_tile(color = "white", size = 0.5) + 
   geom_text(aes(label = N, color = N > max(N)/2), show.legend = FALSE, size = 4) +
  scale_color_manual(values = c("black", "white")) +
  
  # A clean gradient (white to steelblue works well for hydrology reports)
  scale_fill_distiller(palette = "Blues", direction = 1, name = "Count") +  # Ensure all factor levels display on the axes, even if the count is 0
  # Ensure all factor levels display on the axes, even if the count is 0
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  
  labs(
    title = "SUFI model vs Observed Discharge Thresholds",
    x = "Observed Threshold (obs)",
    y = "Modelled Threshold (SUFI)"
  ) +
  
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Removes default grid lines for a cleaner look
    axis.text.x = element_text(angle = 45, hjust = 1) # Angles x-axis labels if they get crowded
)

# ggsave(
#           filename = "../data/figures/qt_heatmap_sim.png",
#           plot = p,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300
#         )

# ggsave(
#           filename = "../data/figures/qt_heatmap_sufi.png",
#           plot = q,
#           width = 8,
#           height = 6,
#           units = "in",
#           dpi = 300
#         )



p + q




lapply(names(comp_dt_list), function(model_config){

  comp_dt <- comp_dt_list[[model_config]]  

  hits_misses <- list()
  for (qt in q_levels){
    hits_misses[[qt]] <- comp_dt[, .(
      Hits         = sum(obs >= qt & mod >= qt, na.rm = TRUE),
      False_Alarms = sum((obs < qt | is.na(obs)) & mod >= qt, na.rm = TRUE),
      Misses       = sum(obs >= qt & (mod < qt | is.na(mod)), na.rm = TRUE),
      n_obs = sum(obs >= qt, na.rm = TRUE)
    ), by = Storm] ## test over qmed
  }


  library(knitr)

  for (qt in q_levels) {
    hits_misses[[qt]][, POD := Hits / (Hits + Misses)]
    hits_misses[[qt]][, FAR := False_Alarms / (Hits + False_Alarms)]
    hits_misses[[qt]][, CSI := Hits / (Hits + Misses + False_Alarms)]
    
    tab <- hits_misses[[qt]][!(Hits == 0 & False_Alarms == 0 & Misses == 0)]    
    # tab <- tab[!POD == "NaN" & !FAR == "NaN" & CSI == "NaN"]
    # cat(paste0(qt, "\n"))
    
    print(
      kable(
        tab,
        caption = paste(model_config, "prediction accuracy metrics for", qt)
      )
    )
  }
})

```

## Peak magnitude error

All sites that are within regions affected for each storm event <br>

Note that the upper scale is cutoff at 98th quantile to prevent outliers controlling colour range. <br>

Grey sites = missing data. <br>

Positive = mod overestimated <br>

```{r echo = FALSE}
#| fig-width: 20
#| fig-height: 10
pme_list <- lapply(events_list, lapply, peak.magnitude.error)

all_pme_plots <- setNames(lapply(names(pme_list), function(mod_config) {
  sim_type <- mod_config

  pme_dt <- data.table(
    unsplit_name = names(unlist(pme_list[[mod_config]])),
    pme = unlist(pme_list[[mod_config]])
  )

  pme_dt[, c("Storm", "G2G.ID") := tstrsplit(unsplit_name, "\\.")]
  pme_dt[, unsplit_name := NULL]
  setcolorder(pme_dt, c("Storm", "G2G.ID", "pme"))

  merged_pme <- merge(pme_dt, qt_dt, by = "G2G.ID")[, .(G2G.ID, Storm, pme, G2G_Easting, G2G_Northing)]

  storm_names <- unique(merged_pme$Storm)

  # Wrap the inner lapply in setNames() so each plot is named by its Storm
  p_list <- setNames(lapply(storm_names, function(name) {
    p <- plot.wales.pme(merged_pme, wales, storm.name = name)
    
    # Optional: Add a subtitle to each plot so you know which model is which when combined
    p <- p + labs(subtitle = paste("Configuration:", sim_type))
    
    return(p)
  }), storm_names)

  # Return the list of plots for this specific model config
  return(p_list) 
  
}), names(pme_list))

all_storms <- names(all_pme_plots$sim)

# Create a list of side-by-side comparisons
comparison_plots <- setNames(lapply(all_storms, function(storm) {
  
  combo_plot <- all_pme_plots$sim[[storm]] + all_pme_plots$sufi[[storm]] +
    plot_annotation(title = paste("Peak Magnitude Error Comparison:", storm))
  
  return(combo_plot)
}), all_storms)

comparison_plots

```

## Extract worst performing sites

```{r}

pme_dt_list <- lapply(pme_list, function(model_config){
  
  pme_dt <- data.table(
    unsplit_name = names(unlist(model_config)),
    pme = unlist(unlist(model_config))
  )

  pme_dt[, c("Storm", "G2G.ID") := tstrsplit(unsplit_name, "\\.")]
  pme_dt[, unsplit_name := NULL]
  setcolorder(pme_dt, c("Storm", "G2G.ID", "pme"))
  merged_pme <- merge(pme_dt, wales_cdata, by = "G2G.ID")[, .(G2G.ID, NRFA.ID, Site.No.,Site.Name., Storm, pme, CATCHMENTSIZE)]
  merged_pme <- merged_pme[complete.cases(merged_pme[,"pme"])] ## remove NA
  return(merged_pme)
})

worst_pme_sim <- pme_dt_list$sim[, .(
  sim_mean_abs_pme = mean((abs(pme)), na.rm = TRUE),
  sim_mean_pme = mean(pme, na.rm = TRUE)), by = .(G2G.ID, Site.No.,Site.Name., NRFA.ID, CATCHMENTSIZE)]

worst_pme_sufi <- pme_dt_list$sufi[, .(
  sufi_mean_abs_pme = mean((abs(pme)), na.rm = TRUE),
  sufi_mean_pme = mean(pme, na.rm = TRUE)), by = .(G2G.ID, Site.No.,Site.Name., NRFA.ID, CATCHMENTSIZE)]


merged_pme <- merge(
  worst_pme_sim,
  worst_pme_sufi,
  by = c("G2G.ID", "Site.No.", "Site.Name.", "NRFA.ID", "CATCHMENTSIZE")
)

merged_pme <- merged_pme[order(sim_mean_pme,sufi_mean_pme,)]

fwrite(merged_pme,"../data/pme_sim_sufi_events_2018_2021.csv")
 

```

<!-- 
TG 135. or a catchment of 605.89 km² (the Teifi at Llandysul is a sizeable, well-known river), a genuine flood peak physically cannot rise and fall that fast — travel time and channel/floodplain storage smooth the hydrograph out over many hours at minimum. A spike that's essentially one or two timesteps wide, with no proportional rise/recession, is a classic signature of a rating-curve extrapolation error or a sensor/logger glitch, not a real flow. High flows are usually converted from a measured water level via a rating curve, and rating curves are often poorly constrained (or not measured at all) far above the range of flows normally gauged — so an unusual stage reading during an extreme event can translate into a wildly wrong discharge estimate. That would explain why the model — which is right about the runoff response — looks like it's dramatically "missing" the peak, when actually the peak itself may not be real (or not that large).
For your worst-negative-PME site list generally, I'd build a small table across all of them capturing: catchment size, KGE/Bias/R² (Sim | SUFI), whether any extreme spike shows the "instant up-and-down" signature, whether it's corroborated by neighbours/rainfall, and whether NRW's QC flags confirm a data issue. That turns "these are the worst sites" into "here's how many of the worst sites are genuine model weaknesses vs. data quality issues" — a much stronger, more defensible conclusion. -->


## Explore impcat of ratings

Note here that rating change means that the observed outputs at those sites were using the old rating curve, not the new one. <br>

```{r echo = FALSE}

# catchment_map <- st_read("P:/NEC04218_Ungauged/releases/ffc/releases/G2G_shapefiles/20250610_g2g_ffc_v2_0")
# wales <- st_transform(wales, 27700) # Ensure CRS is BNG (should already be, but safe to enforce)

rating_change_dt <- fread("../data/rating_change_g2g.csv")
pme_dt_sufi <- pme_dt_list$sufi

pme_dt_sufi_rating <- merge(pme_dt_sufi, rating_change_dt, by = "G2G.ID")

pme_dt_sufi_rating[, rating_group := factor(
  new_rating,
  levels = c(0, 1),
  labels = c("No change in rating curve", "Using old rating")
)]

q1 <- quantile(pme_dt_sufi_rating$pme, 0.25, na.rm = TRUE)
q3 <- quantile(pme_dt_sufi_rating$pme, 0.75, na.rm = TRUE)
iqr <- q3 - q1

pme_clean <- pme_dt_sufi_rating[
  pme >= (q1 - 1.5 * iqr) & pme <= (q3 + 1.5 * iqr)
]
counts <- pme_clean[, .(
  n = .N,
  y = max(pme, na.rm = TRUE)
), by = rating_group]

counts[, y := y + 0.05 * diff(range(pme_clean$pme, na.rm = TRUE))]

ggplot(pme_clean, aes(x = rating_group, y = pme, fill = rating_group)) +
  geom_boxplot(alpha = 0.7) +
  geom_text(
    data = counts,
    aes(x = rating_group, y = y, label = paste0("n = ", n)),
    inherit.aes = FALSE,
    vjust = 0
  ) +
  labs(
    x = "Rating change",
    y = "PME",
    title = "PME by whether site has changed rating since 2022 (SIM)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

pme_dt_sufi <- pme_dt_list$sufi

pme_dt_sufi_rating <- merge(pme_dt_sufi, rating_change_dt, by = "G2G.ID")

pme_dt_sufi_rating[, rating_group := factor(
  new_rating,
  levels = c(0, 1),
  labels = c("No change in rating curve", "Using old rating")
)]

q1 <- quantile(pme_dt_sufi_rating$pme, 0.25, na.rm = TRUE)
q3 <- quantile(pme_dt_sufi_rating$pme, 0.75, na.rm = TRUE)
iqr <- q3 - q1

pme_clean <- pme_dt_sufi_rating[
  pme >= (q1 - 1.5 * iqr) & pme <= (q3 + 1.5 * iqr)
]
counts <- pme_clean[, .(
  n = .N,
  y = max(pme, na.rm = TRUE)
), by = rating_group]

counts[, y := y + 0.05 * diff(range(pme_clean$pme, na.rm = TRUE))]

ggplot(pme_clean, aes(x = rating_group, y = pme, fill = rating_group)) +
  geom_boxplot(alpha = 0.7) +
  # geom_text(
  #   data = counts,
  #   aes(x = rating_group, y = y, label = paste0("n = ", n)),
  #   inherit.aes = FALSE,
  #   vjust = 0
  # ) +
  labs(
    x = "Rating change",
    y = "PME",
    title = "PME by whether site has changed rating since 2022 (SUFI)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```


## Test impact of slope heterogeneity


```{r}

wales_dem <- terra::rast("../data/topography/wales_dem_50m.tif")

slope_rast <- terra::terrain(wales_dem, v = "slope", unit = "degrees")
tri_rast   <- terra::terrain(wales_dem, v = "TRI")

catchments <- st_read("../data/20250610_g2g_ffc_v2_0/fullcats_g2g_ffc_v2_0_loc.shp")

catchments_v <- terra::vect(catchments)

print("Extracting metrics for each catchment...")
slope_extracted <- terra::extract(slope_rast, catchments_v, fun = mean, na.rm = TRUE)
tri_extracted   <- terra::extract(tri_rast, catchments_v, fun = mean, na.rm = TRUE)
tri_range_extracted   <- terra::extract(tri_rast, catchments_v, fun = sd, na.rm = TRUE) ## range might be more interesting for heterogeneity ?

terrain_dt <- data.table(
  G2G.ID     = catchments$g2g_id, 
  Mean_Slope = slope_extracted[[2]],  # The second column holds the actual raster values
  Mean_TRI   = tri_extracted[[2]],
  sd_TRI = tri_range_extracted[[2]]
)

wales_cdata <- merge(wales_cdata, terrain_dt, by = "G2G.ID", all.x = TRUE)

# head(wales_cdata[, .(G2G.ID, River.Name, Mean_Slope, Mean_TRI)])
# merged_pme <- merge(pme_dt_list$sim, pme_dt_list$sufi)

# pme_tri_list <- lapply(pme_dt_list, function(x) merge(x, terrain_dt, by = "G2G.ID"))



# plot(log(abs(sufi_mean_pme)) ~ sd_TRI, data = pme_tri)


```


## River gauge metrics

```{r}

river_gauges <- fread("../data/rg_info_2022.csv")

gauges_sf <- st_as_sf(
  river_gauges, 
  coords = c("X", "Y"), 
  crs = 27700
)

gauges_v <- terra::vect(gauges_sf)

print("Generating proximity-to-gauge distance map...")
gauge_dist_rast <- terra::distance(wales_dem, gauges_v)

print("Extracting mean gauge distance per catchment...")
dist_extracted <- terra::extract(gauge_dist_rast, catchments_v, fun = mean, na.rm = TRUE)
gauge_dist_dt <- data.table(
  G2G.ID     = catchments$g2g_id, 
  mean_dist = dist_extracted[[2]],  
  mean_dist_km = dist_extracted[[2]] /1000 
)


# # pme_dist_tri <- merge(gauge_dist_dt, pme_tri, by = "G2G.ID")
# pme_dist_tri <- lapply(pme_tri_list, function(x) merge(x, gauge_dist_dt, by = "G2G.ID"))


# lapply(pme_dist_tri, function(x){
#   x[, log_abs_pme := log(abs(pme))]
#   x[, Mean_Gauge_Dist_Km := mean_dist / 1000]
# })


# plot(log(abs(sufi_mean_pme)) ~ mean_dist, data = pme_dist)

wales_cdata <- merge(wales_cdata, gauge_dist_dt, by = "G2G.ID", all.x = TRUE)

```


## Model test


- Response variable is pme
- Predictors:
  - catchment size
  - rain gauge distance
  - TRI


```{r}
# merged_pme_predictor <- lapply(pme_dt_list, function(x){
#   merge()
# })


# pme_dist_tri$sufi[,
#     .(Mean_TRI_z := scale(Mean_TRI),
#     mean_dist_z := scale(mean_dist),
#     log_catchment_z := scale(log(CATCHMENTSIZE)))
#   ]
# library(lme4)

# sim_model <- lmer(log_abs_pme ~ Mean_TRI + Mean_Gauge_Dist_Km + log(CATCHMENTSIZE) +
#                  (1 | G2G.ID) + (1 | Storm),
#                data = pme_dist_tri$sim)

# summary(sim_model)

# sufi_model <- lmer(log_abs_pme ~ Mean_TRI + Mean_Gauge_Dist_Km + log(CATCHMENTSIZE) +
#                  (1 | G2G.ID) + (1 | Storm),
#                data = pme_dist_tri$sufi)
# summary(sufi_model)


```


## NRW Threshold exceedance

```{r}


max_discharge_sim_sufi <- lapply(events_list, lapply, extract.peak.discharge)

# nrw_threshold <- fread("../data/nrw_alert_threshold_convert_flow.csv")
# nrw_threshold_v <- nrw_threshold$flow_thresholds_alert
# names(nrw_threshold_v) <- nrw_threshold$G2G.ID
# max_discharge_sim_sufi_filt <- lapply(max_discharge_sim_sufi, lapply, function(x){
#   g2g_sites <- names(nrw_threshold_v)
#   mod <- x$mod
#   obs <- x$obs
#   mod_sites <- gsub("_Mod", "", names(mod))
#   obs_sites <- gsub("_Obs", "", names(obs))
#   filt <- mod_sites %in% g2g_sites
#   mod_filt <- mod[filt]
#   obs_filt <- obs[filt] ## filter so theyre matching
#   return(list(obs = obs_filt, mod = mod_filt))
# })

# nrw_thresh_exceed <- lapply(max_discharge_sim_sufi_filt, lapply, thresh.exceed, threshold = nrw_threshold_v)

# nrw_thresh_exceed_dt_list <- lapply(nrw_thresh_exceed, function(storm){
#     dt_exceed <- data.table(
#     unsplit_name     = names(unlist(storm)),
#     Exceeded = unlist(storm)
#     )
#     dt_exceed[, c("Storm", "Data_Type", "G2G.ID") := tstrsplit(unsplit_name, "\\.", keep = 1:3)]
#     dt_exceed[, G2G.ID := gsub("_Obs|_Mod", "", G2G.ID)]
#     dt_exceed[, unsplit_name := NULL]
#     setcolorder(dt_exceed, c("Storm", "Data_Type", "G2G.ID", "Exceeded"))
#     return(dt_exceed)
# })

# nrw_thresh_exceed_cat <- lapply(nrw_thresh_exceed_dt_list, function(dt) {
#   wide <- dcast(
#     dt,
#     Storm + G2G.ID  ~ Data_Type,
#     value.var = "Exceeded", ## pivot to long with mod and obs filled by exceeded
#     fill = FALSE
#   )

#   wide[, exceed_class := fifelse(obs & mod, "both",
#     fifelse(obs & !mod, "obs only",
#       fifelse(!obs & mod, "mod only", "neither")
#     )
#   )]

#   wide <- na.omit(wide) ## remove NA cols (from obs failing)
#   # wide

#   added_cdata <- merge(wide, wales_cdata, by = "G2G.ID")
# })

# pod_table <- nrw_thresh_exceed_cat$sufi[
#   exceed_class %in% c("both", "obs only", "mod only"),

#   .(
#     hits   = sum(exceed_class == "both"),
#     misses = sum(exceed_class == "obs only"),
#     false_alarms = sum(exceed_class == "mod only")
#   ),

#   by = G2G.ID
# ]

# pod_table[, POD := hits / (hits + misses)]
# pod_table[, FAR := false_alarms / (hits + false_alarms)]


# pod_table <- merge(
#   pod_table,
#   wales_cdata,
#   by = "G2G.ID"
# )

# setorder(pod_table, POD)

# # fwrite(
# #   pod_table,
# #   "../data/nrw_thresh_exceed_pod_by_g2g_site.csv"
# # )

# storm_names <- unique(nrw_thresh_exceed_cat$sim$Storm)

# comparison_plots <- setNames(lapply(storm_names, function(storm_name) {
#   p_sim  <- plot.wales.nrw.thresh(nrw_thresh_exceed_cat$sim,  wales, storm.name = storm_name) +
#     labs(subtitle = "SIM")

#   p_sufi <- plot.wales.nrw.thresh(nrw_thresh_exceed_cat$sufi, wales, storm.name = storm_name) +
#     labs(subtitle = "SUFI")

#   p_sim + p_sufi +
#     plot_annotation(title = paste("Storm:", storm_name))
# }), storm_names)

# comparison_plots



# hits_dt <- lapply(nrw_thresh_exceed_cat, function(dt){
#   dt <- dt[obs == TRUE]
#   dt[, hit := as.integer(mod)]
# })

# far_dt <- lapply(nrw_thresh_exceed_cat, function(dt){
#   dt <- dt[obs == FALSE]
#   dt[, FA := as.integer(mod)]
# })

# ## need to do a podfar
# detection_model_sufi <- glmer(
#   hit ~  log(CATCHMENTSIZE) + mean_dist_km + sd_TRI +
#     (1 | G2G.ID) + (1 | Storm),
#   data = hits_dt$sufi,
#   family = binomial(link = "logit"),
#   control = glmerControl(optimizer = "bobyqa")  # more robust convergence for binomial GLMMs
# )
# summary(detection_model_sufi)




# site_hit_rates <- as.data.table(model.frame(detection_model_sufi))[
#   , .(hit_rate = mean(hit), n = .N), by = G2G.ID]
# table(site_hit_rates$hit_rate %in% c(0,1))
# site_hit_rates[, extreme := hit_rate %in% c(0,1)]

# # does perfect separation cluster among low-n sites?
# site_hit_rates[, .(mean_n = mean(n), min_n = min(n), max_n = max(n)), by = extreme]

# # visualise directly
# ggplot(site_hit_rates, aes(x = n, y = hit_rate, color = extreme)) +
#   geom_jitter(width = 0.1, height = 0.02) +
#   labs(title = "Site hit rate vs number of storms observed")

# # FAR_model_sufi <- glmer(
# #   FA ~  log(CATCHMENTSIZE) + mean_dist_km+sd_TRI +
# #     (1 | G2G.ID) + (1 | Storm),
# #   data = far_dt$sufi,
# #   family = binomial(link = "logit"),
# #   control = glmerControl(optimizer = "bobyqa")  # more robust convergence for binomial GLMMs
# # )
# # summary(FAR_model_sufi)


```


#### With tolerance

```{r}
max_discharge_sim_sufi <- lapply(events_list, lapply, extract.peak.discharge)

nrw_thresh_exceed <- lapply(max_discharge_sim_sufi_filt, lapply, pod.far.tol, threshold = nrw_threshold_v)
nrw_thresh_exceed <- lapply(nrw_thresh_exceed, lapply, merge, y = wales_cdata, by = "G2G.ID")
nrw_thresh_exceed_dt <- lapply(nrw_thresh_exceed, rbindlist, idcol = "Storm")


storm_names <- unique(nrw_thresh_exceed_dt$sim$Storm)

comparison_plots <- setNames(lapply(storm_names, function(storm_name) {
  p_sim  <- plot.wales.nrw.thresh.tol(nrw_thresh_exceed_dt$sim,  wales, storm.name = storm_name) +
    labs(subtitle = "SIM")

  p_sufi <- plot.wales.nrw.thresh.tol(nrw_thresh_exceed_dt$sufi, wales, storm.name = storm_name) +
    labs(subtitle = "SUFI")

  p_sim + p_sufi +
    plot_annotation(title = paste("Storm:", storm_name))
}), storm_names)

comparison_plots

# ## need to do a podfar
# detection_model_sufi <- glmer(
#   hit ~  log(CATCHMENTSIZE) + mean_dist_km + sd_TRI +
#     (1 | G2G.ID) + (1 | Storm),
#   data = hits_dt$sufi,
#   family = binomial(link = "logit"),
#   control = glmerControl(optimizer = "bobyqa")  # more robust convergence for binomial GLMMs
# )
# summary(detection_model_sufi)


detection_data <- lapply(nrw_thresh_exceed_dt, function(dt){
  dt[outcome %in% c("hit", "near miss", "miss")]
})


detection_data <- lapply(detection_data, function(dt){
  dt[, `:=`(
  hit_strict  = as.integer(outcome == "hit"),                    # near miss counts as a miss
  hit_lenient = as.integer(outcome %in% c("hit", "near miss"))   # near miss counts as a hit
)]
})


model_strict <- glmer(hit_strict ~ log(CATCHMENTSIZE) + (1 | G2G.ID) + (1 | Storm),
                       data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

model_lenient <- glmer(hit_lenient ~ log(CATCHMENTSIZE) + (1 | G2G.ID) + (1 | Storm),
                        data = detection_data$sufi, family = binomial, control = glmerControl(optimizer = "bobyqa"))

exp(fixef(model_strict)); exp(fixef(model_lenient))






# site_hit_rates <- as.data.table(model.frame(detection_model_sufi))[
#   , .(hit_rate = mean(hit), n = .N), by = G2G.ID]
# table(site_hit_rates$hit_rate %in% c(0,1))
# site_hit_rates[, extreme := hit_rate %in% c(0,1)]

# # does perfect separation cluster among low-n sites?
# site_hit_rates[, .(mean_n = mean(n), min_n = min(n), max_n = max(n)), by = extreme]

# # visualise directly
# ggplot(site_hit_rates, aes(x = n, y = hit_rate, color = extreme)) +
#   geom_jitter(width = 0.1, height = 0.02) +
#   labs(title = "Site hit rate vs number of storms observed")


```

# things to do 14/07

- see if there are any catchments that perform badly with nrw threshold, pme, and kge
where are they on the map?


# things to do

how are sites performing in general. longer period hydrograph covering few years. KGE statistics at these sites. is that site performing well, or is it the high flows that we're getting.
3 years, across 2018.
Look at NRFA, see if there is anything particularly noticable.


- are there reservoirs
- gauges upstream feeding into it.
- headwater any 
- can we map slopes (tomography)


Also the rainfall that is going in, needs checking.

Raingauges.
- one idea is the amount of raingauges in a certain radius, were g2g site is the centroid.
-  Catchment size split up; altert warnings per catchment
Elwy, putting in thresholds
Gridded outputs.
  - john making gridded outputs
  - can look at different high flows at different parts of the river.
Some of the other storm events, pre 2018.
Do analyses per catchment size.
Pull Ma from midland.
