---
title: "QT Exceedannce plots  for Storm events (Wales 2018-2021)"
subtitle: "Sim data"
author: "Caleb Scutt"
date: "26/06/26"
format:
  html:
    toc: true
    code-fold: true
    fig-width: 12
    fig-height: 8
    theme: flatly
execute:
  warning: false
  message: false
---

# Logic

* Read in G2G data

```{r echo = FALSE}
## load packages and functions
library(data.table)
library(ggplot2)
library(lubridate)
library(gtools)
library(sf)
source("../functions/utility.R")
source("../functions/extract.g2g.qt.R")
source("../functions/qt.comparison.R")
source("../functions/plot.wales.R")

```


## Read in data

* `mod_ts` is the g2g modelled flow data at each g2g site
* `obs_ts` is the observed flow data at each g2g site
* `events` are the notable storm events in Wales (`events_pre_2022` are events we have data for)
* `catchments` are the metadata for each catchment (region, area, river name, easting/northing, g2g.id). `wales_cdata` is this filtered by just Welsh catchments

```{r}
## get g2g simulation file paths (this is Sim)
run_dates <- list.dirs("../data/g2g_data/Sim_g2g_run/", full.names = FALSE, recursive = FALSE)
tsfiles <- paste0(run_dates, "/base_.dat_WA")
tsfiles_filtered <- tsfiles[19:22] ## starts at 2018
tspath <- file.path("..", "data", "g2g_data","Sim_g2g_run")
## read in mod and obs
mod_ts <- read.time.series(tsfiles_filtered,tspath, output = "model")
obs_ts <- read.time.series(tsfiles_filtered,tspath, output = "observed")

## read in the events
events <- fread("../data/notable_events.csv")
events_pre_2022 <- events[Year <= 2021,]
## read in the catchment metadata
catchments <- fread("../data/cdata/catchment_cdata_EA-NRW.csv", fill = Inf)
## filter by wales
wales_cdata <- catchments[Region. == "Wales", ]
```


Here we make a nested list, with top level of list representing each storm event from `events`. \\
We subset Wales into North, Mid and South by these Northing values (can be adjusted). These were used because some sites were not categorised by Area (e.g South East [CY]) \\
We subset the time series 7 days either side of the event beginning/end. \\


```{r}
## classify regions based on northing (can be adjusted)
region_classifier <- list(north_wales = wales_cdata$WISKI.NORTHING >= 300000, 
                          mid_wales = wales_cdata$WISKI.NORTHING >= 230000 & wales_cdata$WISKI.NORTHING <= 300000, 
                          south_wales= wales_cdata$WISKI.NORTHING <= 230000)

## make list for each event, with sublist for mod & obs
events_list <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  region_classifier, exclude.non.notable.sites = FALSE)
```


## Read in QT data

QT data is extracted using functions from `functions/griddedData_class_jrw.py` converted to R.
These functions can be re-used when more catchment data is added.

```{r, message = FALSE}
qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grid_list <- lapply(qt_grid_paths, read.ascii)
names(qt_grid_list) <- qt_val ## name by qt

qt_dt <- make.qt.csv(qt_grid_list, cdata = wales_cdata)

```


## Extract the peak discharge at each flood event

* Peak discharge extracted at each event at each g2g site for both modelled and observed, then compared to each QT.
* If mod / obs discharge is >= QT, TRUE is returned, else FALSE

```{r}
filtered_qt <- lapply(events_list, filter.qt, qt = qt_dt)

max_discharge_qt <- list()
for (t_val in qt_val){
    max_discharge_qt[[t_val]] <- Map(function(event_mod_obs, event_qt){
        extract.peak.discharge(event_mod_obs, event_qt,  t_val)
    }, events_list, filtered_qt)
}


qt_exceeded <- lapply(max_discharge_qt,lapply, compare.qt)

qt_exceeded_dt_list <- lapply(qt_exceeded, function(qt){
    dt_exceed <- data.table(
    path     = names(unlist(qt)),
    Exceeded = unlist(qt)
    )
    dt_exceed[, c("Storm", "Data_Type", "G2G.ID") := tstrsplit(path, "\\.", keep = 1:3)]
    dt_exceed[, G2G.ID := gsub("_Obs|_Mod", "", G2G.ID)]
    dt_exceed[, path := NULL]
    setcolorder(dt_exceed, c("Storm", "Data_Type", "G2G.ID", "Exceeded"))
    return(dt_exceed)
})


merged_dat <- lapply(qt_exceeded_dt_list, merge, qt_dt, by = "G2G.ID")


```


```{r ECHO = FALSE, warning = FALSE, message = FALSE}
wales <- st_read("W:/hymod/Hydro-JULES/HJ Internships/2026 - NRW G2G/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")# Read Wales constituency polygons
wales <- st_transform(wales, 27700) # Ensure CRS is BNG (should already be, but safe to enforce)


flat_dt <- rbindlist(merged_dat, idcol = "Threshold")

exceeded_only <- flat_dt[Exceeded == TRUE]

exceeded_only[, Threshold := gsub("qmed", "q1", Threshold)]

q_levels <- c("q1", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")

exceeded_only[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

setorder(exceeded_only, Storm, G2G.ID, Data_Type, Threshold)
exceeded_only[, Threshold := gsub("\\<q1\\>", "qmed", Threshold)]
q_levels <- c("qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")
exceeded_only[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

highest_qt_exceeded <- unique(
  exceeded_only, 
  by = c("Storm", "G2G.ID", "Data_Type"), 
  fromLast = TRUE
)

```

## Plots

```{r echo = FALSE}

library(gridExtra)



for (name in unique(highest_qt_exceeded$Storm)) {
  p <- plot.wales.exceed(highest_qt_exceeded, wales, storm.name = name, "obs")
  q <- plot.wales.exceed(highest_qt_exceeded, wales, storm.name = name, "mod")
  grid.arrange(p, q, ncol=2)
}

```

## Notable catchments 

```{r}
source("../functions/make.hydrograph.r")
events_list_notable_catchments <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  region_classifier, exclude.non.notable.sites = TRUE)



hydrograph_plots <- list()
for (event in names(events_list_notable_catchments)) {
  event_data <- events_list_notable_catchments[[event]]
    for (g2g_id in gsub("_Mod", "", colnames(event_data$mod)[2:length(colnames(event_data$mod))] )){
       tryCatch(print(make.hydrograph(event_data, g2g_id,storm.name = event, cdata = wales_cdata)), error = function(e) NA)
    }
}




```