---
title: "QT Exceedannce plots  for Storm events (Wales 2018-2021)"
subtitle: "SUFI data"
author: "Caleb Scutt"
format:
  html:
    toc: true
    code-fold: true
    fig-width: 12
    fig-height: 8
    embed-resources: true
    theme: flatly
  typst:
    toc: true
    fig-width: 7
    fig-height: 5
    margin:
      x: 1in
      y: 1in
    fontsize: 11pt
    section-numbering: 1.1.1
    keep-typ: false
execute:
  warning: false
  message: false
  echo: false
---


```{r echo = FALSE}
## load packages and functions
library(data.table)
library(ggplot2)
library(lubridate)
library(gtools)
library(sf)
library(grid)
library(gridExtra)
library(patchwork)
source("../functions/utility.R")
source("../functions/extract.g2g.qt.R")
source("../functions/qt.exceed.R")
source("../functions/plot.wales.R")


```


## Read in data

* `mod_ts` is the g2g modelled flow data at each g2g site
* `obs_ts` is the observed flow data at each g2g site
* `events` are the notable storm events in Wales (`events_pre_2022` are events we have data for)
* `catchments` are the metadata for each catchment (region, area, river name, easting/northing, g2g.id). `wales_cdata` is this filtered by just Welsh catchments
* The following sites are missing from g2g model output, but present in catchment data: 055025_TG_325_Mod, 57809_Mod 60005_Mod, 066004_TG_9512_Mod, 067005_TG_116_Mod, 067023_TG_117_Mod, 055025_TG_325_Mod, 60005_Mod.

```{r, include = FALSE}
## get g2g simulation file paths (this is Sim)
run_dates <- list.dirs("../data/g2g_data/SUFI_g2g_run/", full.names = FALSE, recursive = FALSE)
tsfiles <- paste0(run_dates, "/base_.dat_WA")
tsfiles_filtered <- tsfiles[19:22] ## starts at 2018
tspath <- file.path("..", "data", "g2g_data","SUFI_g2g_run")
## read in mod and obs
mod_ts <- read.time.series(tsfiles_filtered,tspath, output = "model")
obs_ts <- read.time.series(tsfiles_filtered,tspath, output = "observed")

## read in the events
events <- fread("../data/notable_events.csv")
events_pre_2022 <- events[Year <= 2021,]
## read in the catchment metadata
catchments <- fread("../data/cdata/catchment_cdata_EA-NRW.csv", fill = Inf)
## filter by wales
wales_cdata <- catchments[Region. == "Wales", ]
```


Here we make a nested list, with top level of list representing each storm event from `events`. <br>
We subset Wales into North, Mid and South by these Northing values (can be adjusted). These were used because some sites were not categorised by Area (e.g South East [CY]) <br>
We subset the time series 7 days either side of the event beginning/end. <br>


```{r message = FALSE, warning = FALSE, include = FALSE}
## classify regions based on northing (can be adjusted)
region_classifier <- list(north_wales = wales_cdata$WISKI.NORTHING >= 300000, 
                          mid_wales = wales_cdata$WISKI.NORTHING >= 230000 & wales_cdata$WISKI.NORTHING <= 300000, 
                          south_wales= wales_cdata$WISKI.NORTHING <= 230000)

## make list for each event, with sublist for mod & obs
events_list <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  region_classifier, exclude.non.notable.sites = FALSE)
```


## Read in QT data

QT data is extracted using functions from `functions/griddedData_class_jrw.py` converted to R.
These functions can be re-used when more catchment data is added.
Note that these QT values are calculated pre-2022.

```{r, message = FALSE, include = FALSE}
qt_grid_paths <- mixedsort(sort(file.path("../data/qt_grids", list.files("../data/qt_grids")))) ## create relative paths
qt_val <- sub("_.*", "", mixedsort(sort(list.files("../data/qt_grids"))))
qt_grid_list <- lapply(qt_grid_paths, read.ascii)
names(qt_grid_list) <- qt_val ## name by qt

qt_dt <- make.qt.csv(qt_grid_list, cdata = wales_cdata)

```


## Extract the peak discharge at each flood event

* Peak discharge extracted at each event at each g2g site for both modelled and observed, then compared to each QT.
* If mod OR obs discharge is >= QT, TRUE is returned for that column, else FALSE

```{r, include = FALSE}
filtered_qt <- lapply(events_list, filter.qt, qt = qt_dt)

max_discharge_qt <- list()
for (t_val in qt_val){
    max_discharge_qt[[t_val]] <- Map(function(event_mod_obs, event_qt){
        extract.peak.discharge(event_mod_obs, event_qt,  t_val)
    }, events_list, filtered_qt)
}


qt_exceeded <- lapply(max_discharge_qt,lapply, qt.exceed)

qt_exceeded_dt_list <- lapply(qt_exceeded, function(qt){
    dt_exceed <- data.table(
    path     = names(unlist(qt)),
    Exceeded = unlist(qt)
    )
    dt_exceed[, c("Storm", "Data_Type", "G2G.ID") := tstrsplit(path, "\\.", keep = 1:3)]
    dt_exceed[, G2G.ID := gsub("_Obs|_Mod", "", G2G.ID)]
    dt_exceed[, path := NULL]
    setcolorder(dt_exceed, c("Storm", "Data_Type", "G2G.ID", "Exceeded"))
    return(dt_exceed)
})

merged_dat <- lapply(qt_exceeded_dt_list, merge, qt_dt, by = "G2G.ID")

```


```{r ECHO = FALSE, warning = FALSE, message = FALSE, include = FALSE}
wales <- st_read("W:/hymod/Hydro-JULES/HJ Internships/2026 - NRW G2G/SENC_MAY_2026_WA_BFC_-1059615406868623242/SENC_MAY_2026_WA_BFC.shp")# Read Wales constituency polygons
wales <- st_transform(wales, 27700) # Ensure CRS is BNG (should already be, but safe to enforce)


flat_dt <- rbindlist(merged_dat, idcol = "Threshold")

exceeded_only <- flat_dt[Exceeded == TRUE]

exceeded_only[, Threshold := gsub("qmed", "q1", Threshold)]

q_levels <- c("q1", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")

exceeded_only[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

setorder(exceeded_only, Storm, G2G.ID, Data_Type, Threshold)
exceeded_only[, Threshold := gsub("\\<q1\\>", "qmed", Threshold)]
q_levels <- c("qmed", "q5", "q10", "q25", "q50", "q75", "q100", "q200", "q250", "q1000")
exceeded_only[, Threshold := factor(Threshold, levels = q_levels, ordered = TRUE)]

highest_qt_exceeded <- unique(
  exceeded_only, 
  by = c("Storm", "G2G.ID", "Data_Type"), 
  fromLast = TRUE
)

```

## Plots

```{r echo = FALSE}

for (name in unique(highest_qt_exceeded$Storm)) {
  p <- plot.wales.exceed(highest_qt_exceeded, wales, storm.name = name, "obs")
  q <- plot.wales.exceed(highest_qt_exceeded, wales, storm.name = name, "mod")
  
  print(
    (p + q) +
      plot_annotation(title = paste("Storm:", name))
  )
}

```

## Notable catchments 

These are hydrographs taken from the notable catchments highlighted by NRW for each storm event. 


```{r echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
#| fig-width: 15
#| fig-height: 10
source("../functions/make.hydrograph.r")
events_list_notable_catchments <- make.events.list(mod_ts, obs_ts, events_pre_2022, cdata = wales_cdata, region.classifier =  region_classifier, exclude.non.notable.sites = TRUE)

```

```{r echo = FALSE}
#| fig-align: left

library(patchwork)


for (event in names(events_list_notable_catchments)) {

  event_data <- events_list_notable_catchments[[event]]
  event_plots <- list()

  g2g_ids <- gsub("_Mod", "", colnames(event_data$mod)[2:length(colnames(event_data$mod))])

  ord_catchment_size_river <- wales_cdata[G2G.ID %in% g2g_ids][
    order(as.numeric(CATCHMENTSIZE)),
    .SD,
    by = River.Name
  ]

  g2g_ids <- ord_catchment_size_river$G2G.ID

  for (g2g_id in g2g_ids) {
    tryCatch({
      event_plots[[g2g_id]] <- make.hydrograph(
        event.data = event_data,
        g2g.id = g2g_id,
        storm.name = event,
        cdata = wales_cdata,
        qt.data = filtered_qt
      )
    }, error = function(e) {
    })
  }

  event_plots <- event_plots[!sapply(event_plots, is.null)]

  if (length(event_plots) > 0) {

    plot_river <- split(event_plots, ord_catchment_size_river$River.Name)

    for (i in seq_along(plot_river)) {
      river_plot_name <- names(plot_river)[i]
      current_chunk <- plot_river[[i]]

      river_patch <- wrap_plots(current_chunk, ncol = 2) +
        plot_annotation(
          title = paste("Storm:", event, "River:", river_plot_name),
          theme = theme(plot.title = element_text(size = 14, face = "bold"))
        )

      print(river_patch)
    }
  }
}

## perhaps highlight on map next to this where g2g site is?

```

## Hydrographs (obs vs modelled comparison)

### Largest flow mismatches in model vs obs

Hydrographs shown for where there is the largest discrepancies in the QT exceeded (`worst_mismatches$rank_diff>=4`).
Note that the storm events are lumped together here. <br>
Top 10 plotted

```{r echo = FALSE}
#| fig-width: 15
#| fig-height: 20

comp_dt <- dcast(
  highest_qt_exceeded, 
  Storm + G2G.ID + Site + G2G_Easting + G2G_Northing ~ Data_Type, 
  value.var = "Threshold"
) ## pivots table so mod and obs are side by side

comp_dt[, rank_mod := as.numeric(mod)] ## ranked where qmed is 1, q1000 is 10
comp_dt[, rank_obs := as.numeric(obs)]
comp_dt[is.na(rank_mod), rank_mod := 0]
comp_dt[is.na(rank_obs), rank_obs := 0]


comp_dt[, rank_diff := abs(rank_mod - rank_obs)] ## calculate rank differences

worst_mismatches <- comp_dt[!is.na(rank_diff)][order(-rank_diff)]
worst_mismatches <-worst_mismatches[worst_mismatches$rank_diff >=4] ## filter by where qt was a difference of 3


mismatch_plots <- list()
n_mismatch <- min(10, nrow(worst_mismatches))
for (i in 1:n_mismatch) {
  event <- worst_mismatches[i, Storm]
  g2g_id <- worst_mismatches[i, G2G.ID]
  event_data <- events_list[[event]]
  mismatch_plots[[i]] <- tryCatch(make.hydrograph(event_data, g2g_id, storm.name = event, cdata = wales_cdata, qt.data = filtered_qt), error = function(e) NA)
}
mismatch_plots <- mismatch_plots[!sapply(mismatch_plots, is.na)]

wrap_plots(mismatch_plots, ncol = 2) +
  plot_annotation(title = "Largest flow mismatches (rank_diff >= 4)")





```


### False alarms (model exceeds qmed, observed under qmed )

```{r echo = FALSE, warning = FALSE}
#| fig-width: 15
#| fig-height: 20


# model false alarm for flood
model_false_positives <- comp_dt[is.na(obs)]
model_false_positives <- model_false_positives[order(-mod)]

false_positive_plots <- list()
for (i in 1:16){
  event <- model_false_positives[i, Storm]
  g2g_id <- model_false_positives[i, G2G.ID]
  event_data <- events_list[[event]]
  false_positive_plots[[i]] <- tryCatch(make.hydrograph(event_data, g2g_id, storm.name = event, cdata = wales_cdata, qt.data = filtered_qt), error = function(e) NA)
} ## some of these are NA because lack of obs data??

false_positive_plots <- false_positive_plots[!sapply(false_positive_plots, is.na)]

grid.arrange(grobs = false_positive_plots)

```

### Model misses (observed exceeds qmed, model under qmed )

```{r}
#| fig-width: 15
#| fig-height: 20

model_miss <- comp_dt[is.na(mod)]
model_miss <- model_miss[order(-obs)]
model_miss_plots <- list()

for (i in 1:10){
  event <- model_miss[i, Storm]
  g2g_id <- model_miss[i, G2G.ID]
  event_data <- events_list[[event]]
  model_miss_plots[[i]] <- make.hydrograph(event_data, g2g_id, storm.name = event, cdata = wales_cdata, qt.data = filtered_qt)
}
grid.arrange(grobs = model_miss_plots)

```


## Prediction accuracy metrics

* POD = probability of detection (closer to 1 = better)
* FAR = false alarm ratio (closer to 0 = better)
* CSI = critical success ratio  (closer to 1 = better)

```{r echo = FALSE}

event_verification <- comp_dt[, .(
  Hits         = sum(!is.na(mod) & !is.na(obs)),  # 
  False_Alarms = sum(!is.na(mod) & is.na(obs)),   
  Misses       = sum(is.na(mod) & !is.na(obs))    
), by = Storm]

event_verification[, POD := Hits / (Hits + Misses)]                  # Perfect = 1
event_verification[, FAR := False_Alarms / (Hits + False_Alarms)]    # Perfect = 0
event_verification[, CSI := Hits / (Hits + Misses + False_Alarms)]   # Perfect = 1

print(event_verification[order(-CSI)]) ## order by CSI which is most informative score?


```



## Peak magnitude error

```{r}






```
